### 2017-09-14 paper.tex
lots and lots of changes, I think I have a second draft now

% OUTLINE
% =======
%
% 1. Introduction
%
%   Here I will introduce the need for a new high-fidelity, low-frequency map of the sky. The
%   justifications are: 21 cm foreground mapping, galactic structure (?), others?
%
% 2. Observations
%
%   2.1 - Introduce the Owens Valley LWA
%   2.2 - Describe the 100 hour dataset
%   2.3 - Calibration strategy
%   2.4 - Source removal strategy
%   2.5 - Horizon RFI suppression strategy
%   2.6 - Beam fitting work
%   2.7 - Ionospheric conditions
%
% 3. Imaging
%
%   Give a brief overview of other widefield imaging algorithms. This should set the context for the
%   advantages of m-mode analysis.
%
%   3.1 - Introduce m-mode analysis imaging
%   3.2 - Describe my implementation of m-mode analysis on the ASTM
%   3.3 - L-curves (?)
%   3.4 - CLEAN
%
% 4. Results
%
%   Present sky maps with an estimate of the thermal noise in each pixel. Systematic errors will be
%   described in the following sub-sections.
%
%   4.1 - Horizon RFI
%   4.2 - Ionospheric effects
%   4.3 - Beam errors
%   4.4 - Peeling failures
%
% 5. Discussion
%
%   Compare with the Haslam map and Jayce's map.
%
% 6. Conclusion
%
% FIGURES
% =======
%
% * Antenna layout of the OVRO-LWA
% * Before/after snapshot image of the RFI removal
% * Image of the empirical beam (with source tracks)
% * Plot of ionospheric scintillation and refraction of an A-team source
% * Example L-curve
% * Zoom-in of the dirty PSF
% * Sky maps with their boot-strapped thermal noise estimates
% * (Model) image of the horizontal stripes that show up if the horizon RFI is not removed
% * (Model) image of the PSF distortion from ionospheric effects (scintillation, refraction)
% * (Model) image of the systematic errors in the map created from beam errors
% * (Model) image of the errors created from sporadic peeling failures

\documentclass[twocolumn]{aastex61}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\usepackage[noend]{algpseudocode}
\newcommand*\Let[2]{\State #1 $\gets$ #2}
\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

\usepackage{tablefootnote}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}\PackageWarning{TODO:}{#1!}}

\renewcommand{\b}{\pmb}
\newcommand{\tbf}{\textbf}
\renewcommand{\d}{{\rm d}}
\newcommand{\atikh}{\b{\hat a}_\text{Tikhonov}}

\newcommand{\astrocaltech}{
    \affiliation{
        Department of Astronomy,
        California Institute of Technology,
        1200 E California Blvd,
        Pasadena, CA 91125
    }
}

\begin{document}

%\watermark{draft}

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley Long Wavelength Array
}

\author{Michael W. Eastwood}
\astrocaltech

\author{Marin M. Anderson}
\astrocaltech

\author{Ryan M. Monroe}
\astrocaltech

\author{Gregg Hallinan}
\astrocaltech

\author{Other People}
\affiliation{
    At some institution
}

%\author{Lincoln Greenhill}
%\affiliation{Harvard}
%
%\author{Hugh Garsden}
%\affiliation{Harvard}
%
%\author{Stephen Bourke}
%\affiliation{Chalmers}
%
%\author{Esayas B. Shume} % contributed ionosphere text
%\affiliation{Jet Propulsion Laboratory, 4800 Oak Grove Dr, Pasadena, CA 91109}

\correspondingauthor{Michael W. Eastwood}
\email{mweastwood@astro.caltech.edu}

\begin{abstract}
    A host of new low-frequency radio telescopes are attempting to detect the 21~cm transition at
    cosmological distances. These measurements hold promise to directly probe star and galaxy
    formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range they can
    achieve against foreground sources of low-frequency radio emission. Consequently, there is a
    growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for use in
    foreground modeling and removal.  We use the Owens Valley Long Wavelength Array (OVRO-LWA) to
    generate 8 new maps of the sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution and
    at frequencies evenly spaced between 36.528 MHz and 73.152 MHz.  Existing full-sky maps at these
    frequencies have angular resolutions $\ge 2^\circ$ and so these maps are a 10-fold improvement
    in angular resolution.  The thermal noise in each map is $\sim$800 mJy/beam.  These maps are
    generated from the application of a new widefield imaging technique for drift-scanning
    interferometers, Tikhonov-regularized $m$-mode analysis imaging. This algorithm constructs
    images of the entire sky in a single synthesis imaging step with exact treatment of widefield
    effects. We additionally introduce this technique and describe how CLEAN can be adapted to
    deconvolve maps generated by $m$-mode analysis imaging.
\end{abstract}

\keywords{
    cosmology: observations --
    dark ages, reionization, first stars --
    radio continuum: galaxies --
    radio continuum: ISM
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

At redshifts $20 \gtrsim z \gtrsim 7$ the 21~cm hyperfine structure line of neutral hydrogen is
expected to produce a 10 to 100 mK perturbation in the CMB spectrum \citep{2006PhR...433..181F,
2012RPPh...75h6901P}. The amplitude of this perturbation on a given line-of-sight is a function of
the neutral fraction of hydrogen, the baryon overdensity, the spin temperature relative to the CMB
temperature at the given redshift, and the line-of-sight peculiar velocity of the gas.  The spatial
power spectrum of this perturbation is thought to be dominated by inhomogeneous heating of the IGM
at $z\sim 20$ \citep{2014MNRAS.437L..36F}, and by growing ionized bubbles during the EoR at $z\sim
7$ where a detection can constrain the ionizing efficiency of early galaxies, the UV photon
mean-free-path, and the minimum halo mass that can support star formation
\citep{2015MNRAS.449.4246G}.

Current 21~cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21~cm transition
after averaging over the entire sky, and power spectrum experiments that incorporate angular
information to attempt to measure the 3D spatial power spectrum of cosmological 21~cm perturbations.
Ongoing global experiments include EDGES \citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA
\citep{todo_price_2017}, BIGHORNS \citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V},
and SARAS 2 \citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21~cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2008AJ....136..641R} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.5}\,.
\end{equation}
Therefore experiments conservatively need to achieve at least 4 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission (after convolving with the instrumental response) is spectrally smooth. However the
low-frequency radio sky is composed of several components: galactic synchrotron emission, supernova
remnants, radio galaxies, free-free emission and absorption from \ion{H}{2} regions, and a confusing
background of radio sources.  Ideally a foreground removal strategy should be informed by the
measured spatial structure and frequency spectrum of all foreground components.  This possibility is
limited by the availability of suitable high-fidelity low-frequency sky maps on angular scales
ranging from tens of degrees to arcminutes.

The Global Sky Model (GSM) \citep{2008MNRAS.388..247D, 2017MNRAS.464.3486Z} is currently the most
commonly used foreground model. The GSM is a data-driven interpolation of various maps between 10
MHz and 100 GHz. However the majority of information contained in the GSM is derived at frequencies
$>1.4$ GHz where the majority of the modern, high-fidelity input maps are located.  Below 408 MHz,
the GSM is largely driven by the Haslam 408 MHz map \citep{1981A&A...100..209H, 1982A&AS...47....1H,
2015MNRAS.451.4311R}. As a consequence, the GSM is not a faithful representation of the sky below
200 MHz. For instance, \citet{2017MNRAS.469.4537D} report errors of $\pm 50\%$ between the GSM and
their own maps at 74 MHz. This is largely due to the increasing contribution of free-free absorption
and modifications to the synchrotron spectral index at low frequencies.

Recently a host of new low-frequency sky surveys have been conducted including MSSS
\citep{2015A&A...582A.123H}, GLEAM \citep{2015PASA...32...25W}, and TGSS
\citep{2017A&A...598A..78I}. However, the primary data product generated by these surveys is a
catalog of radio point sources. Surveys that capture the diffuse emission include at 45 MHz, where
\citet{2011A&A...525A.138G} produced a map of the sky with 5$^\circ$ resolution, and the LWA1 Low
Frequency Sky Survey \citep{2017MNRAS.469.4537D}, which covers a range of frequencies between 35 MHz
and 80 MHz with resolution betwen 4.5$^\circ$ and 2$^\circ$.

This paper presents a series of new low-frequency maps of the sky between 36.528~Mhz and 73.152~Mhz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21~cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. The maps will be made freely available online at the Legacy Archive for
Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the Owens Valley Long Wavelength Array (OVRO-LWA). In
\S\ref{sec:results} we present the sky maps, and in \S\ref{sec:error} we discuss some of the sources
of error present in the maps. Finally in \S\ref{sec:compare} the new sky maps are compared against
other low-frequency sky maps, and in \S\ref{sec:conclusion} we present our conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{All-Sky Imaging}\label{sec:imaging}

The goal of all imaging algorithms is to estimate the brightness of the sky $I_\nu(\hat r)$ in the
direction $\hat r$ and frequency $\nu$. A radio interferometer measures the visibilities
$V^{ij}_{\nu}$ between pairs of antennas -- here numbered $i$ and $j$ respectively. If the antennas
are separated by the baseline $\vec b_{ij}$, and $A_\nu(\hat r)$ describes their response to the
incident radiation, then
\begin{equation}\label{eq:basic-imaging}
    V^{ij}_\nu = \int_\text{sky}
                 A_\nu(\hat r) I_\nu(\hat r)
                 \exp\bigg(2\pi i \hat r\cdot\vec b_{ij}/\lambda\bigg) \,\d\Omega \, .
\end{equation}
Constructing an image from the output of a radio interferometer consists of estimating $I_\nu(\hat
r)$ given the available measurements $V^{ij}_\nu$.

For later convenience we will define the baseline transfer function $B^{ij}_\nu(\hat r)$ such that
\begin{equation}\label{eq:baseline-transfer-function}
    V^{ij}_\nu = \int_\text{sky} B^{ij}_\nu(\hat r) I_\nu(\hat r) \,\d\Omega \, .
\end{equation}
The baseline transfer function defines the response of a single baseline to the sky, and is a
function of the antenna primary beam, and baseline length and orientation.

Naively one might attempt to solve Equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations, then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the OVRO-LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit complex
floating point number).  This matrix equation is therefore prohibitively large, and solving
Equation~\ref{eq:basic-imaging} by means of discretization is usually impossible although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

Instead it is common to make mild assumptions that simplify Equation~\ref{eq:basic-imaging} and ease
the computational burden in solving for $I_\nu(\hat r)$. For example, when all of the baselines
$\vec b_{ij}$ lie in a plane and the field-of-view is small, Equation~\ref{eq:basic-imaging} can be
well-approximated by a two-dimensional Fourier transform \citep{2001isra.book.....T}. The
restriction on baseline co-planarity and field-of-view can be relaxed by using W-projection
\citep{2008ISTSP...2..647C}. Known primary beam effects can also be accounted for during imaging by
using A-projection \citep{2013ApJ...770...91B}.

\subsection{$m$-Mode Analysis}

On the other hand, transit telescopes can take advantage of a symmetry in
Equation~\ref{eq:basic-imaging} that greatly reduces the amount of computer time required to image
the full-sky with exact incorporation of widefield imaging effects. This technique, called $m$-mode
analysis, also obviates the need for gridding and mosaicing. Instead the entire sky is imaged in one
coherent synthesis imaging step.

In this context we will define a transit telescope as any interferometer where the response pattern
of the individual elements does not change with respect to time. This may be an interferometer like
the OVRO-LWA where the correlation elements are fixed dipoles, but it may also be an interferometer
like LOFAR or the MWA if the steerable beams are held in a fixed position (not necessarily at
zenith).

We will briefly summarize $m$-mode analysis below, but the interested reader should consult
\citet{2014ApJ...781...57S, 2015PhRvD..91h3514S} for a complete derivation.

For a transit telescope, the visibilities $V^{ij}_\nu$ are a periodic function of sidereal
time.\footnote{
    This is not strictly true. Ionospheric fluctuations and non-sidereal sources (such as the sun)
    will violate this assumption. This paper will, however, demonstrate that the impact on the final
    maps is mild.
}
Therefore it is a natural operation to compute the Fourier transform of the visibilities with
respect to sidereal time $\phi\in[0,2\pi)$.
\begin{equation}
    V^{ij}_{m,\nu} = \int_0^{2\pi} V^{ij}_\nu(\phi)\exp\bigg(-im\phi\bigg)\,\d\phi
\end{equation}
The output of this Fourier transform is the set of $m$-modes $V^{ij}_{m,\nu}$ where
$m=0,\,\pm1,\,\pm2,\,\ldots$ is the Fourier conjugate variable to the sidereal time. The $m$-mode
corresponding to $m=0$ is a simple average of the visibilities over sidereal time. Similarly $m=1$
corresponds to the component of the visibilities that varies over half-day timescales. Larger values
of $m$ correspond to components that vary on quicker timescales.

It can be shown that there is a discrete linear relationship between the measured $m$-modes
$V^{ij}_{m,\nu}$ and the spherical harmonic coefficients of the sky brightness $a_{lm,\nu}$.
\begin{equation}\label{eq:m-mode-sum-equation}
    V^{ij}_{m,\nu} = \sum_l B^{ij}_{lm,\nu} a_{lm,\nu}\,,
\end{equation}
where the transfer coefficients $B^{ij}_{lm,\nu}$ define the interferometer's response to the sky.
These transfer coefficients are computed from the spherical harmonic transform of the baseline
transfer function defined by Equation~\ref{eq:baseline-transfer-function}.

Equation~\ref{eq:m-mode-sum-equation} can be recognized as a matrix equation where the transfer
matrix $\b B$ is block-diagonal.
\begin{equation}\label{eq:m-mode-matrix-equation}
    \overbrace{\left(
        \begin{array}{c}
            \vdots \\
            m\text{-modes} \\
            \vdots \\
        \end{array}
    \right)}^{\b v}
    =
    \overbrace{\left(
        \begin{array}{ccc}
            \ddots & & \\
            & \text{transfer matrix} & \\
            & & \ddots \\
        \end{array}
    \right)}^{\b B}
    \overbrace{\left(
        \begin{array}{c}
            \vdots \\
            a_{lm} \\
            \vdots \\
        \end{array}
    \right)}^{\b a}
\end{equation}
\begin{equation}
    \b B = \left(\begin{array}{cccc}
        m = 0 &&& \\
              & m=\pm1 && \\
              && m=\pm2 & \\
              &&& \ddots \\
    \end{array}\right)
\end{equation}
The vector $\b v$ contains the list of $m$-modes and the vector $\b a$ contains the list of
spherical harmonic coefficients representing the sky brightness. In order to take advantage of the
block-diagonal structure in $\b B$, $\b v$ and $\b a$ must be sorted by the absolute value of $m$.
Positive and negative values of $m$ are grouped together because the brightness of the sky is
real-valued, and the spherical harmonic transform of a real-valued function has $a_{l(-m)} = (-1)^m
a_{lm}^*$.

In practice we now need to pick the set of spherical harmonics we will use to represent the sky. For
an interferometer like the OVRO-LWA with lots of short baselines, a sensible choice is to use all
spherical harmonics with $l\le l_\text{max}$ for some $l_\text{max}$. The parameter $l_\text{max}$
is determined by the maximum baseline length of the interferometer.  For an interferometer without
short spacings, a minimum value for $l$ might also be used. This $l_\text{min}$ parameter should be
determined by the minimum baseline length.  When creating the maps presented in this paper, we use
$l_\text{min} = 0$ and $l_\text{max} = 1000$. %\todo{this choice of lmax restricts resolution at
%higher frequencies}

The interferometer's sensitivity to the monopole ($a_{00}$) deserves special consideration.
\citet{2016ApJ...826..116V} prove -- under fairly general assumptions -- that a baseline with
nonzero sensitivity to $a_{00}$ must also have some amount of cross-talk or common-mode noise.  In
fact the sensitivity to $a_{00}$ is proportional to a sum of these effects. For example, one way a
baseline can have nonzero sensitivity to $a_{00}$ is if the baseline is extremely short. In this
case the antennas are so close together that voltage fluctuations in one antenna can couple into the
other antenna. In order to make an interferometric measurement of $a_{00}$, this coupling must be
measured and calibrated. Consequently we set $a_{00}=0$ in our analysis.

The size of a typical block in the transfer matrix is
$(2N_\text{base}N_\text{freq})\times(l_\text{max})$. If each element of the matrix is stored as a
64-bit complex floating point number, a single block is 500 MB for the case of single-channel
imaging with the OVRO-LWA. Compare this number with the 5 PB required for the naive approach.  The
power of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}.  By breaking up the problem into $N$ independent blocks,
the computational complexity involved in inverting a single block is reduced by a factor of $N^3$
relative to a dense matrix. However we must invert $N$ blocks, so the overall computational savings
is proportional to $N^2$.  For the case of the OVRO-LWA the equation breaks up into $\sim10^3$
blocks and so we save a factor of $\sim10^6$ in processing time by using $m$-mode analysis.

\subsection{$m$-Mode Analysis Imaging}

Imaging in $m$-mode analysis essentially amounts to inverting
Equation~\ref{eq:m-mode-matrix-equation} to solve for the spherical harmonic coefficients $\b a$.
The linear-least squares solution, which minimizes $\|\b v - \b B\b a\|^2$, is given by
\begin{equation}
    \b{\hat a}_\text{LLS} = (\b B^*\b B)^{-1}\b B^*\b v\,,
\end{equation}
where $^*$ indicates the conjugate-transpose. However, usually one will find that $\b B$ is not
full-rank and hence $\b B^*\b B$ is not an invertible matrix. For example, an interferometer located
in the northern hemisphere will never see a region of the southern sky centered on the southern
celestial pole. The $m$-modes contained in the vector $\b v$ must contain no information
about the sky around the southern celestial pole, and therefore the act of multiplying by $\b B$
must destroy some information about the sky. The consequence of this fact is that $\b B$ must have
at least one singular value that is equal to zero. It is then trivial to show that $\b B^*\b B$ must
have at least one eigenvalue that is equal to zero, which means it is not an invertible matrix.

Another way of looking at the problem is that because the interferometer is not sensitive to part of
the southern hemisphere, there are infinitely many possible solutions to
Equation~\ref{eq:m-mode-matrix-equation} that will fit the measured data equally well. So we need to
regularize the problem and apply an additional constraint that prefers a unique solution. For
example, you may prefer that in the absence of any information the sky should be set to zero or you
may prefer that the sky should be set to some prior expectation.

\subsubsection{Tikhonov Regularization}

The process of Tikhonov regularization minimizes $\|\b v - \b B\b a\|^2 + \varepsilon\|\b a\|^2$ for
some arbitrary value of $\varepsilon > 0$ chosen by the observer. The solution that minimizes this
expression is given by
\begin{equation}\label{eq:tikhonov-solution}
    \atikh = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b v\,.
\end{equation}
Tikhonov regularization adds a small value $\varepsilon$ to the diagonal of $\b B^*\b B$, fixing the
matrix's singularity. We can see this by using the singular value decomposition (SVD) of the matrix
$\b B = \b U \b \Sigma \b V^*$. Equation~\ref{eq:tikhonov-solution} becomes
\begin{equation}
    \atikh = \b V (\b\Sigma^2 + \varepsilon \b I)^{-1}\b\Sigma \b U^*\b v\,,
\end{equation}
where
\[
    \b\Sigma = \left(
        \begin{array}{ccc}
            \sigma_1 & & \\
                     & \sigma_2 & \\
                     & & \ddots \\
        \end{array}
    \right)\,.
\]
The diagonal elements of $\b\Sigma$ are the singular values of $\b B$ and so the contribution of
each singular component to the Tikhonov-regularized solution is scaled by $\sigma_i / (\sigma_i^2 +
\varepsilon)$, where $\sigma_i$ is the singular value for the $i$th singular component. Tikhonov
regularization therefore acts to suppress any component for which
$\sigma_i\lesssim\sqrt{\varepsilon}$.  If $\sigma_i = 0$, the component is set to zero.

In practice the measurement $\b v$ is corrupted by noise with covariance $\b N$. For illustrative
purposes we will assume that $\b N=n\b I$ for some scalar $n$. In this case the covariance of the
Tikhonov-regularized spherical harmonic coefficients is
\begin{equation}
    \b C = n \b V (\b\Sigma^2 + \varepsilon\b I)^{-2} \b\Sigma^2 \b V^*\,.
\end{equation}
Each singular component is scaled by a factor of $\sigma_i^2/(\sigma_i^2 + \varepsilon)^2$.  In the
absence of Tikhonov regularization ($\varepsilon=0$) singular components with the smallest singular
values -- the ones that the interferometer is the least sensitive to -- actually come to dominate
the covariance of the measured spherical harmonic coefficients. Tikhonov regularization improves
this situation by downweighting these components.

While Tikhonov regularization will force unmeasured modes to zero, if a prior map of the sky already
exists, it will be preferrable to instead minimize $\|\b v - \b B\b a\|^2 + \varepsilon \|\b a-\b
a_\text{prior}\|^2$.

\subsubsection{L-Curves}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/lcurve/lcurve}
    \caption{
        An example L-curve using OVRO-LWA data at 36.528 MHz. The $x$-axis is the norm of the
        solution (in this case the spherical harmonic coefficients) given in arbitrary units. The
        $y$-axis is the least-squares norm given in arbitrary units. $\varepsilon$ is the
        regularization parameter. When the regularization parameter is small, the norm of the
        solution grows rapidly. When the regularization parameter is large, the least-squares norm
        grows rapidly.
    }
    \label{fig:lcurve}
\end{figure}

Tikhonov regularization requires the observer to pick the value of $\varepsilon$. If $\varepsilon$
is too large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. However if $\varepsilon$ is too small then the problem will be
poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\varepsilon$ therefore requires understanding the trade-off between the two norms.

This trade-off can be analyzed quantitatively by trialing several values of $\varepsilon$, and
computing $\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in
Figure~\ref{fig:lcurve}. The shape of this curve has a characteristic L-shape, and as a result this
type of plot is called an L-curve. The ideal value of $\varepsilon$ lies near the turning point of
the plot. At this point a small decrease in $\varepsilon$ will lead to an undesired rapid increase
in $\|\b a\|^2$, and a small increase in $\varepsilon$ will lead to an undesired rapid increase in
$\|\b v - \b B\b a\|^2$.

In practice, the L-curve should be used as a guide to estimate a reasonable value of $\varepsilon$.
However better results can often be obtained by tuning the value of $\varepsilon$. For instance
increasing the value of $\varepsilon$ can improve the noise properties of the map by down-weighting
noisy modes. Decreasing the value of $\varepsilon$ can improve the resolution in the map by
up-weighting the contribution of longer baselines, which are likely fewer in number. In this respect
choosing the value of $\varepsilon$ is analagous to picking the weighting scheme in traditional
imaging where robust weighting schemes can be tuned to similar effect \citep{briggs}.

\subsubsection{The Moore-Penrose Pseudoinverse}

The Moore-Penrose pseudoinverse (denoted in this paper with a superscript $\dagger$), is commonly
applied to find the minimum-norm linear-least squares solution to a set of linear equations. This
can be used in place of Tikhonov regularization as
\begin{equation}
    \b{\hat a}_\text{Moore-Penrose} = \b B^\dagger\b v\,.
\end{equation}
Much like Tikhonov regularization, the Moore-Penrose pseudoinverse sets components with small
singular values (below some user-defined threshold) to zero. Components with large singular values
(above the user-defined threshold) are included in the calculation at their full amplitude with no
down-weighting of modes near the threshold. The essential difference between using the Moore-Penrose
pseudoinverse and Tikhonov regularization is that the pseudoinverse defines a hard transition from
on to off. Modes are either set to zero or included in the map at their full amplitude. On the other
hand Tikhonov regularization smoothly interpolates between these behaviors. Because of this,
Tikhonov regularization tends to produce better results in practical applications.

\subsection{CLEAN}\label{sec:clean}

In traditional radio astronomy imaging, the CLEAN algorithm \citep{1974A&AS...15..417H} is used to
interpolate the measured visibilities onto the $uv$-plane. In the absence of this interpolation,
gaps in the interferometer's $uv$-coverage are assumed to be zero, and -- in the image plane --
sources are convolved with a point spread function (PSF) that is characteristic of the
$uv$-coverage.  Fundamentally, the interferometer's PSF in the dirty images is determined by which
modes were assumed to be zero in the initial imaging process.

In $m$-mode analysis imaging, we assumed modes were zero in two separate ways:
\begin{enumerate}
    \item We selected a set of spherical harmonic coefficients $a_{lm}$ to describe the sky
        brightness distribution. All modes with $l>l_\text{max}$ are neglected and assumed to be
        zero.
    \item Tikhonov regularization forces linear combinations of spherical harmonic coefficients with
        $\sigma_i \lesssim \sqrt{\varepsilon}$ towards zero.
\end{enumerate}
As a consequence the final map of the sky is not assembled from a complete set of spherical
harmonics. Therefore just as in traditional imaging, $m$-mode analysis imaging produces dirty maps
where sources are convolved with a PSF.  This PSF can be improved by increasing the number and
variety of baselines, which increases the number of modes where $\sigma_i \gg \sqrt{\varepsilon}$.
Alternatively by collecting more data, the signal-noise ratio of the measured $m$-modes increases,
which allows the observer to lower the value of $\varepsilon$ without increasing the noise in the
maps.  Finally, the CLEAN algorithm can be applied to interpolate some of the missing information
that was assumed to be zero.

The PSF of a dirty $m$-mode analysis map may be computed with
\begin{equation}\label{eq:psf}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the PSF at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta, \phi)$ is the
vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given by
\begin{equation}
    \b a_\text{PS}(\theta, \phi) = \begin{pmatrix}
        \vdots \\
        Y_{lm}^*(\theta, \phi) \\
        \vdots \\
    \end{pmatrix}
    = \begin{pmatrix}
        \vdots \\
        Y_{lm}^*(\theta, 0)\times e^{im\phi} \\
        \vdots \\
    \end{pmatrix} \,.
\end{equation}
In general the PSF can be a function of the right ascension and declination. However point sources
at the same declination take the same track through the sky and therefore (barring any ionospheric
effects) will have the same PSF. The PSF is therefore only a function of the declination. For
example, sources at low elevations will tend to have an extended PSF along the north-south axis due
to baseline foreshortening. Some examples PSFs are shown in Figure~\ref{fig:psf}.  Adapting CLEAN
for $m$-mode analysis requires either pre-computing Equation~\ref{eq:psf} at a grid of declinations,
or a method for rapidly evaluating Equation~\ref{eq:psf} on the fly.

%\citep{2011MNRAS.413..411B} % I think Lincoln wants me to cite this paper

For an interferometer with more baselines than spherical harmonics used in the maps (eg. the
OVRO-LWA), $\b B^*\b B$ can be a much smaller matrix than the full transfer matrix $\b B$. Therefore
pre-computing $\b B^*\b B$ can allow the entire matrix to fit into memory on a single machine. This
greatly reduces the amount of disk I/O necessary for solving Equation~\ref{eq:psf}.

Additionally we can pre-compute the Cholesky decomposition of $\b B^*\b B + \varepsilon\b I = \b
U^*\b U$, where $\b U$ is an upper-triangular matrix. Inverting an upper triangular matrix is an
$\mathcal{O}(N^2)$ operation (instead of $\mathcal{O}(N^3)$ for a general matrix inverse).\footnote{
    Inverting a matrix is numerically unstable and should be avoided. Instead of computing $\b
    A^{-1}$, the linear equation $\b A\b x = \b b$ is solved each time the matrix inverse is needed.
}
Equation~\ref{eq:psf} can then be rapidly evaluated from right to left as:
\begin{equation}\label{eq:rapid-psf}
    \b a_\text{PSF} =
        \b U^{-1}\,\big(\b U^*\big)^{-1}\,\big(\b B^*\b B\big)\,\b a_\text{PS}\,.
\end{equation}
Furthermore, Equation~\ref{eq:rapid-psf} does not need to be separately evaluated for each CLEAN
component. Instead we can identify $N$ CLEAN components, accumulate $\b a_\text{PS}$ for each
component, and evaluate Equation~\ref{eq:rapid-psf} on the accumulation. This can greatly reduce the
number of times this equation needs to be evaluated, but care must be taken to ensure that the $N$
components are not so close together that sidelobes from one may interact with another.

%\todo{describe blurring to avoid cleaning the diffuse emission}

Altogether the adaptation of CLEAN applied to the maps presented in this paper is summarized below.
\begin{algorithmic}[1]
    \Require{$\b a$ is the solution to Equation~\ref{eq:tikhonov-solution}}
    \Function{CLEAN}{$\b a$}
    \Let{$\b M$}{$\b B^*\b B$}
    \Let{$\b U$}{${\rm chol}(\b M + \varepsilon\b I)$} \Comment{Cholesky decomposition}
    \While{noise in map $>$ threshold}
    \State find $N$ pixels with the largest residual flux
    \Let{$\b x$}{$\sum_{i=1}^N \,(\text{pixel flux}) \times \b a_\text{PS}(\theta_i, \phi_i)$}
    \Let{$\b y$}{$\b U^{-1}\big(\b U^*\big)^{-1}\b M\b x$}
    \Let{$\b a$}{$\b a - (\text{loop gain})\times\b y$}
    \State record subtracted components
    \EndWhile
    \Let{$\b a$}{$\b a + (\text{restored components})$}
    \State \Return{$\b a$}
    \EndFunction
\end{algorithmic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations}\label{sec:observations}

\subsection{The Owens Valley Long Wavelength Array}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/antenna-layout/antenna-layout}
    \caption{
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 triangles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to $\sim1.5$ km. The 5 crosses
        are antennas equipped with noise-switched front ends.
    }
    \label{fig:antenna-layout}
\end{figure}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/psf/psf}
    \caption{
        This figure shows the $m$-mode analysis imaging PSF at $\delta = +45^\circ$ at 3 separate
        frequencies: 36.528 MHz (left panel), 52.224 MHz (middle panel), and 73.152 MHz (right
        panel). The PSF is computed by evaluating Equation~\ref{eq:psf}.
    }
    \label{fig:psf}
\end{figure*}

The Owens Valley Long Wavelength Array (OVRO-LWA) is a 288-element interferometer located at the
Owens Valley Radio Observatory near Big Pine, California \citep{todo_hallinan_2017}.  The OVRO-LWA
is a low-frequency instrument with instantaneous bandwidth covering 27.384 MHz to 84.912 MHz with 24
kHz channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that there are
$288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA correlator
\citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible hemisphere in a
single snapshot image.  In the current configuration 32 antennas (64 signal paths) are unused.

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging.  251 of the antennas are contained within a 200 m diameter core. 32 antennas are
placed outside of the core in order to extend the maximum baseline length to $\sim$1.5~km. The final
5 antennas are equipped with noise-switched front ends for calibrated total power measurements of
the global sky brightness.  These antennas are used as part of the LEDA experiment
\citep{todo_price_2017} to measure the global signal of 21~cm absorption from the cosmic dawn.
Figure~\ref{fig:antenna-layout} is a diagram of the antenna configuration.

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
During this time the OVRO-LWA operated as a zenith-pointing drift scanning interferometer.  The
correlator dump time was selected to be 13 seconds such that the correlator output evenly divides a
sidereal day.

\subsection{Gain Calibration}\label{sec:gaincal}

Antenna gain calibration is accomplished using an iterative method independently developed by
\citet{2008ISTSP...2..707M} and \citet{2014A&A...571A..97S}. The calibration routine is written in
the Julia programming language \citep{doi:10.1137/141000671}, and is publicly available
online\footnote{\url{https://github.com/mweastwood/TTCal.jl}} under an open source license (GPLv3 or
any later version).

Cyg~A and Cas~A are -- by an order of magnitude -- the brightest point-like radio sources in the
northern hemisphere. Therefore the optimal time to solve for the interferometer's gain calibration
is when these sources are at high elevations.  The antenna complex gains are measured from a 22
minute track of data from when Cyg~A and Cas are at high elevations. The gains measured in this way
are then used to calibrate the entire 28 hour dataset. The calibration sky model consists only of
Cyg~A and Cas~A. The model flux of Cyg~A is set to the \citet{1977A&A....61...99B} spectrum while
the flux of Cas~A is measured from the data itself (using a preliminary calibration).  Residual
errors in the astrometry are corrected post-imaging by registering the images with respect to all
VLSSr \citep{2014MNRAS.440..327L} sources brighter than 30~Jy. Additional cuts are made to ensure
that sources aren't too close to each other or other bright sources.

Temperature fluctuations within the electronics shelter generate 0.1 dB sawtooth oscillations in the
analog gain. These oscillations occur with a variable 15 to 17 minute period. The amplitude of these
gain fluctuations is calibrated by smoothing the autocorrelation amplitudes on 45 minute timescales.
The ratio of the measured auto-correlation power to the smoothed auto-correlation power defines a
per-antenna amplitude correction that is then applied to the cross-correlations.

\subsection{Primary Beam Measurements}\label{sec:beam}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/beam/beam}
    \caption{
        This figure shows the empirical fits to the OVRO-LWA Stokes-I primary beam (the response of
        the $x$- and $y$-dipoles has been summed) at three frequencies: 36.528 MHz (left panel),
        52.224 MHz (middle panel), and 73.152 MHz (right panel). The source tracks used to measure
        the beam model are overlaid. From north to south, these tracks correspond to Cas~A, Cyg~A,
        Per~B, Tau~A, Vir~A, Her~A, 3C~353, and Hya~A.  The fitting process is described in
        \S\ref{sec:beam}, and residuals for Cyg~A and Cas~A can be seen in
        Figure~\ref{fig:scintillation}.
    }
    \label{fig:beam}
\end{figure*}

In order to generate wide-field images of the sky, the response of the antenna to the sky must be
known. Drift-scanning interferometers like the OVRO-LWA can empirically measure their
primary beam under a mild set of symmetry assumptions \citep{2012AJ....143...53P}. In this work we
assume that the primary beam is invariant under north-south and east-west flips, and additionally
that the $x$- and $y$-dipoles have the same response to the sky after rotating one by 90$^\circ$.
These symmetries are apparent in the antenna design, but real-world defects and coupling with nearby
antennas will contribute towards breaking these symmetries at some level. However some amount of
symmetry must be assumed in order to break the degeneracy between source flux and beam amplitude
when the flux of a source is unknown.

We measure the flux of several bright sources (Cyg~A, Cas~A, Tau~A, Vir~A, Her~A, Hya~A, Per~B, and
3C~353) as they pass through the sky and then fit a beam model composed of Zernike polynomials to
those flux measurements. We select the basis functions to have the desired symmetry ($Z_0^0$,
$Z_2^0$, $Z_4^0$, $Z_4^4$, $Z_6^0$, $Z_6^4$, $Z_8^0$, $Z_8^4$, and $Z_8^8$) and the beam amplitude
at zenith is constrained to be unity. See Figure~\ref{fig:beam} for an illustration of a fitted beam
model at several frequencies. This process is repeated for each frequency channel.

\subsection{Ionospheric Conditions}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/vtec/vtec}
    \caption{
        The median vertical TEC within 200 km of OVRO during the time of the observation. The gray
        shaded regions indicate times outside of the observing period. The gray vertical lines
        indicate sunrise and sunset (as labeled).
    }
    \label{fig:vtec}
\end{figure}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/scintillation-refraction/scintillation-refraction}
    \caption{
        \textbf{(a, b)} These panels show the measured apparent flux of Cyg~A and Cas~A at 36.528
        MHz (red points) and 73.152 MHz (blue points) as a function of time over the observing
        period. The solid black curves show the expected flux computed using the empirical beam
        model fits. The thermal noise contribution to each point is about 50 Jy. The majority of the
        scatter is due to ionospheric scintillation.  Cyg~A is occulted by the White Mountains when
        it is low on the horizon to the east.
        \textbf{(c, d)} These panels show the measured position offset of Cyg~A and Cas~A relative
        to their true astronomical positions at 36.528 MHz (red line) and 73.152 MHz (blue line).
    }
    \label{fig:scintillation}
\end{figure*}

The geomagnetic conditions during this time were mild. The Disturbance storm time (Dst) index was
$>-30$ nT during the entirety of the observing period.\footnote{
    The Dst index was obtained from the World Data Center for Geomagnetism, Kyoto University
    (\url{http://swdcwww.kugi.kyoto-u.ac.jp/}). Accessed 2017-07-25.
}
Following the classification scheme of \citet{2008GMS...181.....K}, a weak geomagnetic storm has
$\text{Dst} < -30$ nT. Stronger geomagnetic storms have $\text{Dst} \ll -30$ nT.

Despite the mild conditions, low frequency interferometric observations are still affected by the
index of refraction in the ionosphere.  Figure~\ref{fig:vtec} shows the median vertical total
electron content (TEC) above OVRO measured from GPS \citep{1999JASTP..61.1205I}. The median is
computed over all GPS measurements within 200 km of the observatory. Over the observing period the
TEC smoothly varies from 20 TECU at midday to 5 TECU during the night. However, this measurement is
only sensitive to large scale fluctuations in the ionosphere and doesn't capture small scale
fluctuations.

Small scale fluctuations are best characterized by source scintillation and refractive offsets.
Figure~\ref{fig:scintillation} shows the apparent flux and position offset of Cyg~A and Cas~A as a
function of time over the entire observing period. Both sources exhibit rapid scintillation on the
timescale of a single integration (13 seconds). For example at 36.528 MHz, it is not unusual for Cyg
A to have a measured flux of 30,000 Jy in one integration, and 20,000 Jy in the next. The measured
position offset of each source is a measurement of the ionospheric phase gradient across the array.
This varies on slower 10 minute timescales refracting each source by as much as 20 arcmin (at 36.528
MHz) from their true astronomical positions as waves in the ionosphere pass through the line of
sight. The impact of these effects on the sky maps is simulated in \S\ref{sec:ionosphere}.

\subsection{Source Removal}

\subsubsection{Cygnus A and Cassiopeia A}

Without removing bright sources from the data, sidelobes from bright sources will dominate the
variance in the sky maps.  At 74 MHz Cyg~A is a 15,000~Jy source \citep{1977A&A....61...99B}. A
conservative estimate for the confusion limit at 74 MHz with a 15~arcmin beam is 1000~mJy
\citep{2014MNRAS.440..327L}. Therefore we require that Cyg~A's sidelobes be at least $-45$~dB down
from the main lobe to prevent Cyg~A's sidelobes from dominating the variance in the image.

At low frequencies, propagation effects through the ionosphere must be accounted for in order to
achieve high dynamic range images. This necessitates the use of direction-dependent calibration and
peeling \citep{2008ISTSP...2..707M, 2015MNRAS.449.2668S}.  In the dataset used in this paper,
scintillation and refractive-offset events on the timescale of a single integration (13 seconds) are
observed (see Figure~\ref{fig:scintillation}. Therefore the direction dependent calibration changes
on these timescales and the we must solve for one set of complex gains per source per integration.

The largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular scale of Cyg~A is
$\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines, the OVRO-LWA marginally
resolves both sources. A resolved source model is needed for both sources. We fit a self-consistent
resolved source model to each source. This is performed by minimizing the variance within an
aperture located on each source after peeling. By phasing up a large number of integrations before
imaging (at least 1 hour) it is possible to smear out the contribution of the rest of the sky.  We
then use NLopt's Sbplx routine \citep{nlopt, sbplx} to vary the parameters in a source model until
the variance within the aperture is minimized.  Armed with resolved source models for both Cyg~A and
Cas~A, these two sources can be peeled from the dataset to the required dynamic range.

\subsubsection{Other Bright Sources}

Other bright sources namely Vir~A, Tau~A, Her~A, Hya~A, Per~B, and 3C~353 are also removed from
the visibilities prior to imaging. Because these sources are much fainter than Cyg~A and Cas~A, we
do not need resolved source models to be able to remove these sources from the visibilities without
residual sidelobes contaminating the image.

However, the ionosphere will cause these sources to scintillate and refract. So the position and
flux of each source is measured separately in each channel and integration. The sources are then
subtracted from the visibilities using the updated position and flux of the source. The brightest of
these sources (Vir~A and Tau~A) are peeled using a direction dependent calibration when they are
at high elevations.

\subsubsection{The Sun}

The sun can be trivially removed from any map of the sky by constructing the map using only data
taken at night. A map of the entire sky can be obtained by using observations spaced 6 months apart.
However the dataset used in this paper consists of 28 consecutive hours. Therefore we attempt to
suppress the contribution of the sun to the maps with the goal of suppressing its sidelobes.  The
sun is well-resolved by the OVRO-LWA and hence a detailed source model is needed. In fact the
optical depth $\tau=1$ surface of the sun changes with frequency and as a consequence a new model is
needed for each of the 8 channel maps produced in this paper. While we could fit a limited number of
Gaussian components to Cyg~A and Cas~A, this is insufficient for the Sun.  Additionally while most
astronomical sources at these frequencies have negative spectral indices, the Sun has a positive
spectral index. Therefore more care will need to be taken in subtracting the Sun at higher
frequencies than at lower frequencies.

The strategy used for removing the Sun below 55 MHz involves fitting a shapelet
\citep{2003MNRAS.338...35R} model to the Sun and subtracting without the use of direction dependent
gains. The shapelet fitting is performed in the visibility space. Above 55 MHz a model is fit to the
Sun by minimizing the residuals after peeling (in the same way that models are obtained for Cyg~A
and Cas~A). The Sun is then peeled from each integration using direction dependent gains.

\subsection{Flux Scale}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/flux-scale/flux-scale}
    \caption{
        This figure shows the measured fluxes (black points) of 11 sources plotted against the
        published spectra from \citet{2017ApJS..230....7P} (solid line above 50 MHz, dotted line
        below 50 MHz), \citet{2012MNRAS.423L..30S} (dashed line), and \citet{1977A&A....61...99B}
        (dot-dash line).
    }
    \label{fig:flux-scale}
\end{figure*}

The flux scale of the data was tied to the \citet{1977A&A....61...99B} spectrum of Cyg~A during gain
calibration. However gain calibration is also a function of the beam model and the spectrum used for
Cas~A. Recent work by \citet{2012MNRAS.423L..30S} (hereafter SH12) using archival date from the
literature and \citet{2017ApJS..230....7P} (hereafter PB17) using the VLA has expanded the number of
low-frequency radio sources with calibrated flux measurements from one (just Cyg~A) to eleven in
total. While the SH12 flux scale is valid between 30 MHz and 300 MHz, the PB17 flux scale is
somewhat more limited because the lowest frequency observations come from the VLA 4-band system. As
a consequence the PB17 flux scale is not valid below 50 MHz.

Figure~\ref{fig:flux-scale} shows a comparison between flux measurements made using the all-sky maps
from this work, and spectra from the aforementioned flux scales. Generally the OVRO-LWA flux
measurements agree to between 5 and 10\% with the SH12 spectra. Below 50 MHz there can be
substantial departures with respect to the extrapolated PB17 spectra (for eg.  3C 286, 3C 295, and
3C 380), but it is usually the case that we have much better agreement with the SH12 spectra. This
indicates that the PB17 spectra cannot be extrapolated below 50 MHz.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:results}

\begin{figure*}[ht]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw04} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw06} \\
    \end{tabular}
    \caption{
        These 8 panels illustrate (with a Mollweide projection and logarithmic color scale) the 8
        full-sky maps generated with Tikhonov-regularized $m$-mode analysis imaging and the
        OVRO-LWA.  Each map covers the sky north of $\delta=-30^\circ$ with angular resolution of
        $\sim15$ arcmin. Eight bright sources have been removed from the map (Cyg~A, Cas~A, Vir~A,
        Tau~A, Her~A, Hya~A, Per~B, and 3C~353). The small blank region near $l=+45.7^\circ$,
        $b=-47.9^\circ$ corresponds to the location of the Sun during the observation period.  A
        detailed summary of the properties of each map is given in Table~\ref{tab:summary}.
    }
    \label{fig:channel-maps}
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw08} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw10} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw12} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw14} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw16} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw18} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/spectral-index/internal-spectral-index}
    \caption{
        This Mollweide-projected map is constructed from 3 maps of the sky at 36.528~MHz (red),
        52.224~MHz (green), and 73.152~MHz (blue). The maps are scaled by $\nu^{2.5}$ before
        combining and the color scale is logarithmic (as in Figure~\ref{fig:channel-maps}). The
        local spectral index (estimated by means of a local T-T plot) is overlaid as a contour plot.
        \todo{
            The spectral index near the galactic plane is very wrong because the local T-T plot is
            contaminated with multiple types of emission. I need to redo this plot by replacing the
            galactic plane with hand-picked regions for estimating the spectral index.
        }
    }
    \label{fig:internal-spectral-index}
\end{figure*}

\begin{table*}[t]
    \centering
    \begin{tabular}{ccccccccc}
        \hline
        \hline
        & \tbf{Frequency} & \tbf{Bandwidth} & \tbf{FWHM}\footnote{
                The full-width half-maximum (FWHM) of the synthesized beam at $\delta=+45^\circ$.
            }&
            \multicolumn2c{\tbf{Thermal Noise}\footnote{
                Measured with a jackknife and splitting the dataset into even- and odd-numbered
                integrations.
            }} &&
            \multicolumn2c{\tbf{Singular Values}\footnote{
                Singular values of the transfer matrix compared with the value of the regularization
                parameter $\varepsilon$ used while solving Equation~\ref{eq:tikhonov-solution}. As
                discussed in the text, singular vectors with corresponding singular values $\sigma
                \ll \sqrt{\varepsilon}$ are set to zero by the Tikhonov regularization procedure.
            }} \\
        \cline{5-6}
        \cline{8-9}
        \tbf{\#} & MHz & MHz & arcmin &
            K & mJy/beam &&
            fraction $>\sqrt{\varepsilon}$ &
            fraction $>\sqrt{10\varepsilon}$ \\
        \hline
        1 & 36.528 & 0.024 & 18.5 & 595. & 799. && 0.391 & 0.077 \\
        2 & 41.760 & 0.024 & 17.2 & 541. & 824. && 0.480 & 0.070 \\
        3 & 46.992 & 0.024 & 16.3 & 417. & 717. && 0.504 & 0.056 \\
        4 & 52.224 & 0.024 & 15.6 & 418. & 814. && 0.535 & 0.053 \\
        5 & 57.456 & 0.024 & 15.4 & 354. & 819. && 0.542 & 0.049 \\
        6 & 62.688 & 0.024 & 15.3 & 309. & 843. && 0.540 & 0.046 \\
        7 & 67.920 & 0.024 & 15.3 & 281. & 894. && 0.529 & 0.047 \\
        8 & 73.152 & 0.024 & 15.7 & 154. & 598. && 0.512 & 0.051 \\
        \hline \hline
    \end{tabular}
    \caption{A summary of the generated all-sky maps}
    \label{tab:summary}
\end{table*}

Eight sky maps were constructed using Tikhonov-regularized $m$-mode analysis imaging and cleaning
with observations from the OVRO-LWA. These maps are shown in Figure~\ref{fig:channel-maps}. Each map
covers the sky north of $\delta=-30^\circ$ with $\sim 15$ arcmin angular resolution resolution. The
8 brightest northern hemisphere point sources are removed from each map (Cyg~A, Cas~A, Vir~A, Tau~A,
Her~A, Hya~A, Per~B, and 3C~353). There is a small blank region near $l=+45.7^\circ$,
$b=-47.9^\circ$ corresponding to the position of the Sun during the observing window. The properties
of each map, including frequency, bandwidth, angular resolution, and thermal noise are presented in
Table~\ref{tab:summary}.

Due to the considerations presented by \citet{2016ApJ...826..116V}, each of these maps is monopole
subtracted. Consequently a simple pixel-by-pixel powerlaw fit will be unduly effected by the choice
of zero-level correction. Instead, temperature-temperature plots (T-T plot) can be used to measure
the spectral index independently of any zero-level corrections \citep{1962MNRAS.124..297T}. This
method relies on the assumption that all pixels in a given region are described by the same power
law. In Figure~\ref{fig:internal-spectral-index} the spectral index is locally estimated within a
region $\sim10^\circ$ across by constructing a local T-T plot at every pixel between the maps at
36.528~MHz and 73.152~MHz. The spectral index contours are overlaid on a 3-color image constructed
from the maps at 36.528~MHz, 52.224~MHz, and 73.152~MHz where the increasing contribution of
free-free absorption in the galactic plane is apparent as blue regions. Similarly the spectral index
can be seen to steepen as galactic latitude increases (this is a transition from blue to white to
yellow that is also evident in the spectral index contours).

In order to suppress sources of terrestrial interference, all spherical harmonics with $m=0$, or
$m=1$ and $l>100$ are filtered from the map (where the spherical harmonics are defined in the J2017
coordinate system). As will be discussed in \S\ref{sec:rfi}, these spherical harmonics are
particularly susceptible to contamination by RFI and common-mode problems. As a consequence,
astronomical emission that circles the J2017 north celestial pole (NCP) is filtered from the maps.
This tends to create rings around the NCP at the declination of bright point sources. These rings
are naturally removed from the map during CLEANing as long as this filtering step is including in
the PSF calculation.

The thermal noise in each map is empirically measured using jackknife resampling. The dataset is
split into even- and odd-numbered integrations. These two groups are imaged and cleaned
independently, before being compared against the maps constructed from all of the available data
using the standard jackknife variance estimator. This estimate of the variance includes all sources
of error that operate on $\sim13$ second timescales (the integration time), but does not account for
more slowly varying effects (for example, sidereal variation in the system temperature or day-night
fluctuations in the ionosphere). These thermal noise calculations are summarized in
Table~\ref{tab:summary}.  VLSSr source counts \citep{2014MNRAS.440..327L} suggest that the confusion
limit at 74 MHz and 15 arcmin angular resolution is $\sim 1000\times(\nu/74\,{\rm MHz})^{-0.7}$ mJy.
Each channel map achieves thermal noise $<900$ mJy and therefore each map is likely at or near the
confusion limit.

Each map from Figure~\ref{fig:channel-maps} is freely available online in Healpix format
\citep{2005ApJ...622..759G} on LAMBDA (\url{https://lambda.gsfc.nasa.gov/}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Error Analysis}\label{sec:error}

\subsection{Terrestrial Interference and Pickup}\label{sec:rfi}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/terrestrial-interference/smeared}
    \caption{
        This figure illustrates terrestrial sources of correlated noise that are apparent after
        averaging the visibilities at 62.688 MHz over the entire 28 hour observing period (keeping
        the phase center at zenith such that astronomical sources of radio emission are smeared
        along tracks of constant declination). Each panel represents a different component that is
        removed from the visibilities. The images are generated using WSCLEAN
        \citep{2014MNRAS.444..606O}, uniform weighting, and only baselines longer than 15
        wavelengths. Panels (a) and (b) illustrate components that appear noise-like in image-space,
        but are in fact a constant-offset to the measured visibilities likely associated with
        cross-talk or common-mode pickup. Panel (c) illustrates a component that is clearly
        associated with an RFI source on the horizon to the west-north-west of the OVRO-LWA. This
        RFI source is likely an arcing power line.  Figure~\ref{fig:rings} illustrates the
        characteristic ring-like artifacts introduced into the maps if these 3 components are not
        removed prior to $m$-mode analysis imaging.
    }
    \label{fig:fitrfi}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/terrestrial-interference/rings}
    \caption{
        This is a Mollweide-projected image of the artifacts introduced to the $m$-mode analysis
        maps by the three terrestrial sources shown in Figure~\ref{fig:rings}. Because these sources
        are not moving through the sky sidereally, they tend to be smeared along rings of constant
        declination. The spurs seemingly radiating from the north celestial pole are a Moir\'{e}
        pattern (ie. an artifact of the pixelization).
    }
    \label{fig:rings}
\end{figure*}

When writing down Equation~\ref{eq:basic-imaging}, it is implicitly assumed that the correlated
voltage fluctuations measured between pairs of antennas are exclusively generated by astronomical
sources of radio emission. In practice this implicit assumption can be violated. For instance a
low-frequency interferometer located in the vicinity of an arcing power line will see an additional
contribution from the radio-frequency interference (RFI) generated by the arcing process. Similarly
common-mode pickup along the analog signal path of the interferometer may generate an additional
spurious contribution to the measured visibilities. While the amplitude and phase of these
contaminating signals may fluctuate with time, they do not sweep across the sky at the sidereal rate
characteristic of astronomical sources.

The Owens Valley is an important source of water and power for the city of Los Angeles.
Unfortunately this means that high voltage power lines run along the valley to the west of the OVRO-
LWA. Some of these power line poles have faulty insulators that arc and produce pulsed, broadband
RFI. Because these poles exist in the near-field of the array, we have been able to localize some of
them by using the curvature of the incoming wavefront to infer a distance. Efforts are currently
underway to work with the utility pole owners to have these insulators replaced.

In the meantime it is possible to suppress their contamination in the dataset. The contribution of
these RFI sources to the visibilities can be plainly seen by averaging $>24$ hours of data with the
phase center set to zenith. In this way, true sky components are smeared along tracks of constant
declination while terrestrial sources (ie. the arcing power lines or any contribution due to
common-mode pickup) are not smeared.  Obtaining a model for the RFI is complicated by the fact that
the contaminating sources are at extremely low elevations where the antenna response is essentially
unknown (and inhomogeneous due to antenna shadowing effects). Therefore it is not enough to know the
physical location of the faulty insulator generating the RFI. In addition you must also know the
response of each antenna (amplitude and phase) in the appropriate direction. This motivates the use
of peeling, which allows the antenna response to be a free parameter.  Therefore model visibilities
for the RFI can be obtained by peeling the sources after smearing the visibilities over $>24$ hours.
See Figure~\ref{fig:fitrfi} for an illustration of some of the removed components at 62.688 MHz.

While attempting to peel RFI sources from the averaged visibilities, it was discovered that
frequently peeling would prefer to remove components from the visibilities that are not obviously
associated with any source on the horizon or elsewhere in the sky (see panels (a) and (b) in
Figure~\ref{fig:fitrfi}).  These components appear noise-like in the images, but they are actually a
constant offset to the measured visibilities and are therefore likely associated with cross-talk or
some form of common-mode pickup. If these components are not subtracted from the measured
visibilities, they contribute ring-like structures into the sky maps as seen in
Figure~\ref{fig:rings}. This figure is not a simulation, but rather a difference between maps
created before and after measuring and subtracting the components in Figure~\ref{fig:fitrfi} from
each integration.

The first step in Equation~\ref{eq:tikhonov-solution} is to compute $\b B^*\b v$. In this step we
compute the projection of the measurement $\b v$ onto the space spanned by the columns of $\b B$.
Each column of $\b B$ describes the interferometer's response to a corresponding spherical harmonic
coefficient of the sky brightness distribution. Therefore the act of computing $\b B^*\b v$ is to
project the measured $m$-modes onto the space of $m$-modes which could be generated by astronomical
sources. The degree to which a source of terrestrial interferer will contaminate a map generated
using $m$-mode analysis imaging is determined by its amplitude after projection.

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = \b 0$.
In practice this is unlikely. In general the contamination is proportional to the overall amplitude
of the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/\|\b v_\text{terrestrial}\|$).

\todo{Compute B*v for Figure~\ref{fig:fitrfi} to show that the pick-up is pretty bad. My expectation
is that it is actually worse than the RFI, but I don't actually know for sure}

These terrestrial sources do not rotate with the sky and hence their contamination tends to be
restricted to modes with small $m$. In this dataset the contamination is largely restricted to $m
\lesssim 1$. Although the RFI is capable of fluctuating on short timescales, in practice the
artifacts it introduces seems to be restricted to small $m$ (presumably because the phase is not
fluctuating).  As a result if the contamination is not suppressed, it will manifest itself as rings
along stripes of constant declination. This effect is plainly visible in Figure~\ref{fig:fitrfi}.
Because of the distinctive ring-like pattern created by terrestrial sources, we chose to discard
spherical harmonics with either $m=0$ or $m=1$ and $l>100$ in order to further suppress the
contamination.

\subsection{Beam Errors}

A good model of the antenna beam is essential for widefield imaging. Because $m$-mode analysis
imaging operates on a full sidereal day of data, images are constructed after watching each point in
the sky move through a large slice of the beam (excepting the celestial poles). The beam model
therefore serves two purposes:

\begin{enumerate}
    \item setting the flux scale as a function of declination
    \item reconciling observations from two separate sidereal times
\end{enumerate}

In the first case, all sources at a given declination take the same path through the antenna primary
beam. If the antenna response is overestimated along this track then all sources at this declination
will have their flux underestimated. Similarly if the antenna response is underestimated then all
the sources will have their flux overestimated. The errors in Figure~\ref{fig:flux-scale} do not
show a clear pattern with declination. Two sources show a clear systematic error at all frequencies:
3C 353 and 3C 380. 3C 353 is the second southernmost source, but Hya A -- the first southernmost
source -- does not exhibit this systematic error. Similarly 3C 380 is at a comparable declination to
Lyn A, which appears if anything to have its flux overestimated. The absence of a coherent pattern
doesn't eliminate the possibility of beam errors effecting the flux scale, but it does mean that
these errors are at least comparable to the errors inherent to the flux scale itself.

The second case is more subtle. Sources are observed at a wide range of locations in the primary
beam of the antenna. The imaging process must reconcile all of these observations together. The beam
model essentially provides the instructions on how to do this. For example if at a time $t_1$ the
antenna gain towards a source is $g$ and at a later time $t_2$ the antenna gain towards that same
source (which has now moved) is $g/2$, then in order to correctly estimate the flux of the source
the observations from time $t_1$ need to be multiplied by $1/g$ but the observations from time $t_2$
need to be multiplied by $2/g$. A mistake in the beam model here will lead to a mistake in the
estimated flux.  However in general the amplitude and phase of the antenna response needs to be
known and so beam model errors can lead to observations being combined with incorrect phase as well.
This will lead to a degraded PSF in the final map. This effect is negligible compared to the
ionospheric effects discussed in \S\ref{sec:ionosphere}.

\subsection{The Ionosphere}\label{sec:ionosphere}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/ionospheric-simulations/ionospheric-simulations}
    \caption{
        This figure illustrates the corrupting influence of the ionosphere at 36.528 MHz. Each panel
        shows the simulated PSF at the position of Cas~A. The left panel shows the PSF in the event
        that there is no ionospheric scintillation and no refractive offsets. This is the PSF used
        to deconvolve the maps with CLEAN. The middle panel shows the PSF with the measured
        ionospheric scintillation for Cas~A from Figure~\ref{fig:scintillation} but no refractive
        offsets. The right panel shows the PSF with the measured refractive offsets for Cas~A from
        Figure~\ref{fig:scintillation} but no scintillation. This demonstrates that the long
        synthesis images generated by $m$-mode analysis imaging are relatively insensitive to
        scintillation, but uncorrected refractive offsets can limit the achievable dynamic range.
    }
    \label{fig:ionospheric-simulations}
\end{figure*}

One of the key assumptions made by $m$-mode analysis is that the sky is completely static.  We
assume that the only time-dependent behavior is the rotation of the Earth, which slowly rotates the
sky through the fringe patterns of the interferometer. At low frequencies the ionosphere violates
this assumption. In particular, ionospheric scintillation will cause even static sources to exhibit
significant variability (see Figure~\ref{fig:scintillation}).

The correlation observed on a given baseline for a single point source is
\begin{equation}
    V_\nu(t_{\textrm{sidereal}}) = I_\nu B_\nu(t_{\textrm{sidereal}}),
\end{equation}
where $I_\nu$ is the flux of the source at the frequency $\nu$, and $B_\nu$ is the baseline transfer
function defined by Equation~\ref{eq:baseline-transfer-function}. The transfer function is a
function of the direction to the source, which is in turn a function of the sidereal time
$t_{\textrm{sidereal}}$. If the source is varying, from intrinsic variability or due to
scintillation, than the source flux is also a function of the time coordinate $t$ such that
\begin{equation}
    V_\nu(t_{\textrm{sidereal}}) = I_\nu(t) B_\nu(t_{\textrm{sidereal}}),
\end{equation}
where $t_{\textrm{sidereal}} = (t \mod 23.9345\,\textrm{hours})$.

In order to compute the $m$-modes we must Fourier transform with respect to the sidereal time. As a
consequence of the Fourier convolution theorem we find:
\begin{equation}\label{eq:no-longer-block-diagonal}
    V_{\nu, m} \sim \sum_{m^\prime} V_{m^\prime}^\textrm{static} I_{\nu, m-m^\prime}\,.
\end{equation}
where $V_{\nu, m}^{\textrm{static}}$ is the set of observed $m$-modes if the source was actually
static, and $I_{\nu, m-m^\prime}$ is the Fourier transform of the light curve $I_{\nu}(t)$.
Equation~\ref{eq:no-longer-block-diagonal} indicates that power is scattered between different
values of $m$. As a consequence, the true transfer matrix, which is exactly block diagonal in the
ideal case, is no longer truly block diagonal \citep{richard_ionosphere_thoughts}.

The maps presented in Figure~\ref{fig:channel-maps} do not account for any off-diagonal terms
arising from ionospheric fluctuations. The effect of this can be seen in
Figure~\ref{fig:ionospheric-simulations}. In this simulation a point source is placed at the
location of Cas A. In the first panel the source flux traces the beam model exactly and the source
is always located exactly at the location of Cas A. In the middle panel, the source is allowed to
scintillate in the same way Cas A does in Figure~\ref{fig:scintillation}. The image of the point
source is effectively unchanged from the ideal PSF. In the right panel, the source position is
allowed to vary in the same way Cas A does in Figure~\ref{fig:scintillation}. The refractive offsets
lead to a clear degradation of the sidelobes. Because the sidelobes are altered from that of the
ideal PSF, refractive offsets will restrict the dynamic range it is possible to obtain with the
CLEAN algorithm described in \S\ref{sec:clean}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparisons with Other Sky Maps}\label{sec:compare}

\subsection{Guzm\'{a}n 45 MHz Map}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/guzman/guzman}
    \caption{
        This Mollweide projection map compares the fractional difference between the Guzm\'{a}n 45
        MHz map, and the OVRO-LWA map at 46.992 MHz (degraded to $5^\circ$ resolution). A positive
        value indicates regions where the OVRO-LWA map has more emission that the Guzm\'{a}n map,
        and a negative value indicates regions where the Guzm\'{a}n map has more emission than the
        OVRO-LWA map. The red patch around the north celestial pole corresponds to where the
        Guzm\'{a}n map has a blank region.  The larger blue patch around the south celestial pole
        corresponds to where the OVRO-LWA map has a blank region. Note also that the OVRO-LWA map
        has several bright point sources subtracted that are not subtracted from the Guzm\'{a}n map.
    }
    \label{fig:guzman-comparison}
\end{figure*}

A direct comparison between the OVRO-LWA maps interpolated to 45 MHz and the Guzm\'{a}n 45 MHz map
\citep{2011A&A...525A.138G} can be seen in Figure~\ref{fig:guzman-comparison}. In order to make this
comparison the OVRO-LWA map was degraded to a 5$^\circ$ resolution by convolving with a Gaussian
kernel, and the Guzm\'{a}n map has had spherical harmonics with $m=0$ discarded in order to make it
consistent with the maps presented in this paper. This figure shows a $\sim20\%$ excess of emission
in the galactic plane that is consistent with the discrepancy observed by the LWA1 sky survey
\citep{2017MNRAS.469.4537D}. However, while the LWA1 sky survey has an excess of emission near the
north galactic pole, no such excess is observed in this work. Instead there is a 10\% excess of
emission near the south galactic pole. Elsewhere off the plane of the galaxy the discrepancy can be
as much as $-20\%$.

\subsection{Haslam 408 MHz Map}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/haslam/haslam-spectral-index}
    \caption{
        The spectral index between the reprocessed Haslam 408 MHz map and the 73.152 MHz map
        (degraded to 56 arcmin resolution). The white regions of the plot correspond to areas where
        there was a negative correlation between the brightness temperature at 73.152 MHz and at 408
        MHz. This can either be due to the presence of multiple emission mechanisms in close
        proximity to each other or due to insufficient dynamic range in temperature leading to
        effectively no correlation between the two temperatures. Regardless, a single power law is
        insufficient to describe the data in these areas.
        \todo{
            Cas A is the extremely blue feature in this plot. It needs to be blotted out because
            it's not present in my maps, but it is present in the Haslam maps. For whatever reason
            this isn't a problem with Cyg A.
        }
    }
    \label{fig:haslam-spectral-index}
\end{figure*}

The Haslam 408 MHz map \citep{1981A&A...100..209H, 1982A&AS...47....1H} was reprocessed by
\citet{2015MNRAS.451.4311R} to remove artifacts associated with $1/f$ noise and bright sources. In
order to compare the OVRO-LWA maps with the lower resolution Haslam map, the OVRO-LWA maps were
degraded to 56 arcmin resolution and the spectral index between the 73.152 MHz map and the 408 MHz
Haslam map was computed by means of a T-T plot at each pixel in the maps (in order to obviate the
need for a zero-level correction to the 73.152 MHz map). The results of this computation are shown
in Figure~\ref{fig:haslam-spectral-index}. Because this comparison uses a local T-T plot to estimate
the spectral index, the angular resolution of this computation is somewhat degraded relative to
other comparisons in the literature but the spectral indices in
Figure~\ref{fig:haslam-spectral-index} are robust to errors in the zero-level in either map.

The primary difference between the spectral index derived between 73.152 MHz and 408 MHz
(Figure~\ref{fig:haslam-spectral-index}), and the spectral index derived between 36.528 MHz and
73.152 MHz (Figure~\ref{fig:internal-spectral-index}) is the appearance of a multitude of \ion{H}{2}
regions that dot the galactic plane in the latter. The free-free absorption from these \ion{H}{2}
regions contributes to an overall flattening of the spectral index in the galactic plane that is
also seen at higher frequencies.

%\subsection{The Global Sky Model}?
%\subsection{LWA1 Sky Survey}?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:conclusion}

In this work we presented a new imaging technique -- Tikhonov regularized $m$-mode analysis imaging
and cleaning -- for drift-scanning telescopes like the OVRO-LWA.  This technique exactly corrects
for widefield effects in interferometric imaging, but offers a $\sim10^6$ times speed-up relative to
the naive approach for the OVRO-LWA. We applied this technique to a 28 hour dataset and generated 8
sky maps between 36.528~MHz and 73.152~MHz. These sky maps are a substantial improvement over
existing maps at these frequencies due to their 15~arcmin angular resolution and $<600$~K thermal
noise. The point source flux scale is consistent with that defined by \citet{2012MNRAS.423L..30S} to
about 5\% and large angular scales are consistent with the work of \citet{2011A&A...525A.138G} to
within 20\%.  These maps are primarily intended to be used as part of a foreground modeling and
subtraction routine for 21~cm cosmology experiments.

At frequencies above $\sim55$~MHz, the angular resolution of these maps is limited by the selection
of $l_\text{max}=1000$. Future work will increase $l_\text{max}$ to remove this restriction, as well
as include more time and bandwidth to improve the thermal noise. The usage of nighttime only data
can help mitigate dynamic range limitations from the ionosphere and also prevent the Sun from
showing up in the maps. Observations could also be extended to slightly higher and lower frequencies
($\sim27$ to $85$~MHz) to take advantage of the full frequency range of the OVRO-LWA. The higher
frequencies are particularly interesting in order to maximize the overlap with the MWA in the
southern hemisphere, which could be used to fill-in the hole around the southern celestial pole.

\bibliographystyle{apj}
\bibliography{paper}

\end{document}