### 2017-10-04 paper.tex
big set of changes trying to address some of joe's comments

% OUTLINE
% =======
%
% 1. Introduction
%
%   Here I will introduce the need for a new high-fidelity, low-frequency map of the sky. The
%   justifications are: 21 cm foreground mapping, galactic structure (?), others?
%
% 2. Observations
%
%   2.1 - Introduce the Owens Valley LWA
%   2.2 - Describe the 100 hour dataset
%   2.3 - Calibration strategy
%   2.4 - Source removal strategy
%   2.5 - Horizon RFI suppression strategy
%   2.6 - Beam fitting work
%   2.7 - Ionospheric conditions
%
% 3. Imaging
%
%   Give a brief overview of other widefield imaging algorithms. This should set the context for the
%   advantages of m-mode analysis.
%
%   3.1 - Introduce m-mode analysis imaging
%   3.2 - Describe my implementation of m-mode analysis on the ASTM
%   3.3 - L-curves (?)
%   3.4 - CLEAN
%
% 4. Results
%
%   Present sky maps with an estimate of the thermal noise in each pixel. Systematic errors will be
%   described in the following sub-sections.
%
%   4.1 - Horizon RFI
%   4.2 - Ionospheric effects
%   4.3 - Beam errors
%   4.4 - Peeling failures
%
% 5. Discussion
%
%   Compare with the Haslam map and Jayce's map.
%
% 6. Conclusion
%
% FIGURES
% =======
%
% * Antenna layout of the OVRO-LWA
% * Before/after snapshot image of the RFI removal
% * Image of the empirical beam (with source tracks)
% * Plot of ionospheric scintillation and refraction of an A-team source
% * Example L-curve
% * Zoom-in of the dirty PSF
% * Sky maps with their boot-strapped thermal noise estimates
% * (Model) image of the horizontal stripes that show up if the horizon RFI is not removed
% * (Model) image of the PSF distortion from ionospheric effects (scintillation, refraction)
% * (Model) image of the systematic errors in the map created from beam errors
% * (Model) image of the errors created from sporadic peeling failures

\documentclass[twocolumn]{aastex61}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\usepackage[noend]{algpseudocode}
\newcommand*\Let[2]{\State #1 $\gets$ #2}
\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

\usepackage{tablefootnote}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}\PackageWarning{TODO:}{#1!}}

\renewcommand{\b}{\pmb}
\newcommand{\tbf}{\textbf}
\renewcommand{\d}{{\rm d}}
\newcommand{\atikh}{\b{\hat a}_\text{Tikhonov}}

\newcommand{\astrocaltech}{
    \affiliation{
        Department of Astronomy,
        California Institute of Technology,
        1200 E California Blvd,
        Pasadena, CA 91125
    }
}

\begin{document}

%\watermark{draft}

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley Long Wavelength Array
}

\author{Michael W. Eastwood}
\astrocaltech

\author{Marin M. Anderson}
\astrocaltech

\author{Ryan M. Monroe}
\astrocaltech

\author{Gregg Hallinan}
\astrocaltech

\author{Other People}
\affiliation{
    At some institution
}

%\author{Lincoln Greenhill}
%\affiliation{Harvard}
%
%\author{Hugh Garsden}
%\affiliation{Harvard}
%
%\author{Stephen Bourke}
%\affiliation{Chalmers}
%
%\author{Esayas B. Shume} % contributed ionosphere text
%\affiliation{Jet Propulsion Laboratory, 4800 Oak Grove Dr, Pasadena, CA 91109}

\correspondingauthor{Michael W. Eastwood}
\email{mweastwood@astro.caltech.edu}

\begin{abstract}
    A host of new low-frequency radio telescopes are attempting to detect the 21~cm transition at
    cosmological distances. These measurements hold promise to directly probe star and galaxy
    formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range they can
    achieve against foreground sources of low-frequency radio emission. Consequently, there is a
    growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for use in
    foreground modeling and removal.  We use the Owens Valley Long Wavelength Array (OVRO-LWA) to
    generate 8 new maps of the sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution and
    at frequencies evenly spaced between 36.528~MHz and 73.152~MHz with 24~kHz bandwidth.  Existing
    full-sky maps at these frequencies have angular resolutions $\ge 2^\circ$ and so these maps are
    a 10-fold improvement in angular resolution.  The thermal noise in each map is $\sim$800
    mJy/beam.  Future iterations will incorporate total power radiometry, and additional bandwidth
    and integration time. This will improve the thermal noise while simultaneously pushing to higher
    angular resolutions with the planned expansion of the OVRO-LWA to 2.6~km baselines. These maps
    are generated from the application of a new widefield imaging technique for drift-scanning
    interferometers, Tikhonov-regularized $m$-mode analysis imaging. This algorithm constructs
    images of the entire sky in a single synthesis imaging step with exact treatment of widefield
    effects. We additionally introduce this technique and describe how CLEAN can be adapted to
    deconvolve maps generated by $m$-mode analysis imaging.  These maps serve as a first step on the
    path to the use of more sophisticated foreground filters in 21~cm cosmology incorporating the
    measured angular and frequency structure of all foreground contaminants.
\end{abstract}

\keywords{
    cosmology: observations --
    dark ages, reionization, first stars --
    radio continuum: galaxies --
    radio continuum: ISM
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

At redshifts $20 \gtrsim z \gtrsim 7$ the 21~cm hyperfine structure line of neutral hydrogen is
expected to produce a 10 to 100 mK perturbation in the Cosmic Microwave Background (CMB) spectrum
\citep{2006PhR...433..181F, 2012RPPh...75h6901P}. The amplitude of this perturbation on a given
line-of-sight is a function of the neutral fraction of hydrogen, the baryon overdensity, the spin
temperature relative to the CMB temperature at the given redshift, and the line-of-sight peculiar
velocity of the gas.  The spatial power spectrum of this perturbation is thought to be dominated by
inhomogeneous heating of the IGM at $z\sim 20$ \citep{2014MNRAS.437L..36F}, and by growing ionized
bubbles during the EoR at $z\sim 7$ where a detection can constrain the ionizing efficiency of early
galaxies, the UV photon mean-free-path, and the minimum halo mass that can support star formation
\citep{2015MNRAS.449.4246G}.

Current 21~cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21~cm transition
after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21~cm perturbations.  Ongoing global experiments include EDGES
\citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA \citep{2017arXiv170909313P}, BIGHORNS
\citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2
\citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21~cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2008AJ....136..641R} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.5}\,.
\end{equation}
Therefore experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. However the low-frequency radio sky is composed of several
components: galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission
and absorption from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally a
foreground removal strategy should be informed by the measured spatial structure and frequency
spectrum of all foreground components.  This possibility is limited by the availability of suitable
high-fidelity low-frequency sky maps on angular scales ranging from tens of degrees to arcminutes.

The Global Sky Model (GSM) \citep{2008MNRAS.388..247D, 2017MNRAS.464.3486Z} is currently the most
commonly used foreground model. The GSM is a data-driven interpolation of various maps between 10
MHz and 100 GHz. However the majority of information contained in the GSM is derived at frequencies
$>1.4$ GHz where the majority of the modern, high-fidelity input maps are located.  Below 408 MHz,
the GSM is a strong function of the Haslam 408 MHz map \citep{1981A&A...100..209H,
1982A&AS...47....1H, 2015MNRAS.451.4311R}. As a consequence, the GSM is not a faithful
representation of the sky below 200 MHz. For instance, \citet{2017MNRAS.469.4537D} report errors of
$\pm 50\%$ between the GSM and their own maps at 74 MHz. This is largely due to the increasing
contribution of free-free absorption and modifications to the synchrotron spectral index at low
frequencies.

Recently a host of new low-frequency sky surveys have been conducted including MSSS
\citep{2015A&A...582A.123H}, GLEAM \citep{2015PASA...32...25W}, and TGSS
\citep{2017A&A...598A..78I}. However, the primary data product generated by these surveys is a
catalog of radio point sources. At 45 MHz, \citet{2011A&A...525A.138G} created a map of the sky that
captures the diffuse emission with 5$^\circ$ resolution.  The LWA1 Low Frequency Sky Survey
\citep{2017MNRAS.469.4537D} similarly maps the sky at a range of frequencies between 35~MHz and
80~MHz with resolution between 4.5$^\circ$ and 2$^\circ$.

This paper presents a series of new low-frequency maps of the sky between 36.528~MHz and 73.152~MHz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21~cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. We anticipate additional anciliary science taking advantage of the combination
of high fidelity and high resolution of these maps, including but not limited to studies of the
cosmic ray emissivity at low frequencies, searches for giant radio galaxies, and constraining the
galactic synchrotron spectrum. The maps will be made freely available online at the Legacy Archive
for Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the Owens Valley Long Wavelength Array (OVRO-LWA). In
\S\ref{sec:results} we present the sky maps, and in \S\ref{sec:error} we discuss some of the sources
of error present in the maps. Finally in \S\ref{sec:compare} the new sky maps are compared against
other low-frequency sky maps, and in \S\ref{sec:conclusion} we present our conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{All-Sky Imaging}\label{sec:imaging}

The goal of all imaging algorithms is to estimate the brightness of the sky $I_\nu(\hat r)$ in the
direction $\hat r$ and frequency $\nu$.  A radio interferometer measures the visibilities
$V^{ij,pq}_{\nu}$ between pairs of antennas numbered $i$ and $j$ respectively, and between
polarizations labeled $p$ and $q$ respectively. We will neglect subtleties associated with polarized
imaging, so the Stokes-$I$ visibilities are constructed from the sum of the $xx$ and $yy$
correlations such that $V^{ij}_{\nu} = (V^{ij,xx}_{\nu}+V^{ij,yy}_{\nu})/2$.  If the antennas are
separated by the baseline $\vec b_{ij}$, and $A_\nu(\hat r)$ describes an antenna's response to the
incident Stokes-$I$ radiation (here assumed to be the same for each antenna), then
\begin{equation}\label{eq:basic-imaging}
    V^{ij}_\nu = \int_\text{sky}
                 A_\nu(\hat r) I_\nu(\hat r)
                 \exp\bigg(2\pi i \hat r\cdot\vec b_{ij}/\lambda\bigg) \,\d\Omega \, .
\end{equation}
Constructing an image from the output of a radio interferometer consists of estimating $I_\nu(\hat
r)$ given the available measurements $V^{ij}_\nu$.

For later convenience we will define the baseline transfer function $B^{ij}_\nu(\hat r)$ such that
\begin{equation}\label{eq:baseline-transfer-function}
    V^{ij}_\nu = \int_\text{sky} B^{ij}_\nu(\hat r) I_\nu(\hat r) \,\d\Omega \, .
\end{equation}
The baseline transfer function defines the response of a single baseline to the sky, and is a
function of the antenna primary beam, and baseline length and orientation.

Naively one might attempt to solve Equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations, then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the OVRO-LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit complex
floating point number).  This matrix equation is therefore prohibitively large, and solving
Equation~\ref{eq:basic-imaging} by means of discretization is usually impossible although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

Instead it is common to make mild assumptions that simplify Equation~\ref{eq:basic-imaging} and ease
the computational burden in solving for $I_\nu(\hat r)$. For example, when all of the baselines
$\vec b_{ij}$ lie in a plane and the field-of-view is small, Equation~\ref{eq:basic-imaging} can be
well-approximated by a two-dimensional Fourier transform \citep{2001isra.book.....T}. The
restriction on baseline co-planarity and field-of-view can be relaxed by using W-projection
\citep{2008ISTSP...2..647C}. Known primary beam effects can also be accounted for during imaging by
using A-projection \citep{2013ApJ...770...91B}.

\subsection{$m$-Mode Analysis}\label{sec:mmode-analysis}

On the other hand, transit telescopes can take advantage of a symmetry in
Equation~\ref{eq:basic-imaging} that greatly reduces the amount of computer time required to image
the full-sky with exact incorporation of widefield imaging effects. This technique, called $m$-mode
analysis, also obviates the need for gridding and mosaicing. Instead the entire sky is imaged in one
coherent synthesis imaging step.  We will briefly summarize $m$-mode analysis below, but the
interested reader should consult \citet{2014ApJ...781...57S, 2015PhRvD..91h3514S} for a complete
derivation.

In the context of $m$-mode analysis a transit telescope is any interferometer for which the response
pattern of the individual elements does not change with respect to time. This may be an
interferometer like the OVRO-LWA where the correlation elements are fixed dipoles, but it may also
be an interferometer like LOFAR or the MWA if the steerable beams are held in a fixed position (not
necessarily at zenith). The interferometer also does not necessarily have to be homogeneous.
Heterogeneous arrays composed of several different types of antennas are allowed as long as care is
taken to generalize Equation~\ref{eq:basic-imaging} for a heterogeneous array.

For a transit telescope, the visibilities $V^{ij}_\nu$ are a periodic function of sidereal
time.\footnote{
    This is not strictly true. Ionospheric fluctuations and non-sidereal sources (such as the Sun)
    will violate this assumption. This paper will, however, demonstrate that the impact on the final
    maps is mild.
}
Therefore it is a natural operation to compute the Fourier transform of the visibilities with
respect to sidereal time $\phi\in[0,2\pi)$.
\begin{equation}
    V^{ij}_{m,\nu} = \int_0^{2\pi} V^{ij}_\nu(\phi)\exp\bigg(-im\phi\bigg)\,\d\phi
\end{equation}
The output of this Fourier transform is the set of $m$-modes $V^{ij}_{m,\nu}$ where
$m=0,\,\pm1,\,\pm2,\,\ldots$ is the Fourier conjugate variable to the sidereal time. The $m$-mode
corresponding to $m=0$ is a simple average of the visibilities over sidereal time. Similarly $m=1$
corresponds to the component of the visibilities that varies over half-day timescales. Larger values
of $m$ correspond to components that vary on quicker timescales.

\citet{2014ApJ...781...57S, 2015PhRvD..91h3514S} show that there is a discrete linear relationship
between the measured $m$-modes $V^{ij}_{m,\nu}$ and the spherical harmonic coefficients of the sky
brightness $a_{lm,\nu}$.
\begin{equation}\label{eq:m-mode-sum-equation}
    V^{ij}_{m,\nu} = \sum_l B^{ij}_{lm,\nu} a_{lm,\nu}\,,
\end{equation}
where the transfer coefficients $B^{ij}_{lm,\nu}$ are computed from the spherical harmonic transform
of the baseline transfer function defined by Equation~\ref{eq:baseline-transfer-function}. These
transfer coefficients define the interferometers response to the corresponding spherical harmonic
coefficients.

Equation~\ref{eq:m-mode-sum-equation} can be recognized as a matrix equation where the transfer
matrix $\b B$ is block-diagonal.
\begin{equation}\label{eq:m-mode-matrix-equation}
    \overbrace{\left(
        \begin{array}{c}
            \vdots \\
            m\text{-modes} \\
            \vdots \\
        \end{array}
    \right)}^{\b v}
    =
    \overbrace{\left(
        \begin{array}{ccc}
            \ddots & & \\
            & \text{transfer matrix} & \\
            & & \ddots \\
        \end{array}
    \right)}^{\b B}
    \overbrace{\left(
        \begin{array}{c}
            \vdots \\
            a_{lm} \\
            \vdots \\
        \end{array}
    \right)}^{\b a}
\end{equation}
\begin{equation}
    \b B = \left(\begin{array}{cccc}
        m = 0 &&& \\
              & m=\pm1 && \\
              && m=\pm2 & \\
              &&& \ddots \\
    \end{array}\right)
\end{equation}
The vector $\b v$ contains the list of $m$-modes and the vector $\b a$ contains the list of
spherical harmonic coefficients representing the sky brightness. In order to take advantage of the
block-diagonal structure in $\b B$, $\b v$ and $\b a$ must be sorted by the absolute value of $m$.
Positive and negative values of $m$ are grouped together because the brightness of the sky is
real-valued, and the spherical harmonic transform of a real-valued function has $a_{l(-m)} = (-1)^m
a_{lm}^*$.

In practice we now need to pick the set of spherical harmonics we will use to represent the sky. For
an interferometer like the OVRO-LWA with many short baselines, a sensible choice is to use all
spherical harmonics with $l\le l_\text{max}$ for some $l_\text{max}$. The parameter $l_\text{max}$
is determined by the maximum baseline length of the interferometer.  For an interferometer without
short spacings, a minimum value for $l$ might also be used. This $l_\text{min}$ parameter should be
determined by the minimum baseline length. A rough estimate of $l$ for a baseline of length $b$ at
frequency $\nu$ is $l \sim \pi b\nu/c$. Based on this estimate for the OVRO-LWA and other
computational considerations we therefore adapt $l_\text{min}=1$ and $l_\text{max}=1000$ across all
frequencies. However this choice of $l_\text{max}$ actually limits the angular resolution above
55~MHz, and therefore future work will increase $l_\text{max}$ to obtain better angular resolution.

The interferometer's sensitivity to the monopole ($a_{00}$) deserves special consideration.
\citet{2016ApJ...826..116V} prove -- under fairly general assumptions -- that a baseline with
nonzero sensitivity to $a_{00}$ must also have some amount of cross-talk or common-mode noise.  In
fact the sensitivity to $a_{00}$ is proportional to a sum of these effects. For example, one way a
baseline can have nonzero sensitivity to $a_{00}$ is if the baseline is extremely short. In this
case the antennas are so close together that voltage fluctuations in one antenna can couple into the
other antenna. In order to make an interferometric measurement of $a_{00}$, this coupling must be
measured and calibrated. Consequently we set $a_{00}=0$ in our analysis. In the future this
limitation will be addressed with the inclusion of calibrated total power radiometry.

The size of a typical block in the transfer matrix is
$(2N_\text{base}N_\text{freq})\times(l_\text{max})$. If each element of the matrix is stored as a
64-bit complex floating point number, a single block is 500 MB for the case of single-channel
imaging with the OVRO-LWA, which a modern computer can easily store and manipulate in memory.
However, with additional bandwidth these blocks quickly become unwieldy and so as a first pass the
analysis in this paper is restricted to single-channel imaging. Note also that for the OVRO-LWA
$N_\text{base} \gg l_\text{max}$, so there are more measurements than unknowns in
Equation~\ref{eq:m-mode-matrix-equation}.

The power of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$.  By splitting the
equation into $N$ independent blocks, the number of floating point operations required to operate on
the full matrix is reduced by a factor of $N^2$. This computational savings is what makes this
matrix algebra approach to interferometric imaging feasible.

\subsection{$m$-Mode Analysis Imaging}\label{sec:mmode-imaging}

Imaging in $m$-mode analysis essentially amounts to inverting
Equation~\ref{eq:m-mode-matrix-equation} to solve for the spherical harmonic coefficients $\b a$.
The linear-least squares solution, which minimizes $\|\b v - \b B\b a\|^2$, is given by
\begin{equation}
    \b{\hat a}_\text{LLS} = (\b B^*\b B)^{-1}\b B^*\b v\,,
\end{equation}
where $^*$ indicates the conjugate-transpose.

However, usually one will find that $\b B$ is not full-rank and hence $\b B^*\b B$ is not an
invertible matrix. For example, an interferometer located in the northern hemisphere will never see
a region of the southern sky centered on the southern celestial pole. The $m$-modes contained in the
vector $\b v$ must contain no information about the sky around the southern celestial pole, and
therefore the act of multiplying by $\b B$ must destroy some information about the sky. The
consequence of this fact is that $\b B$ must have at least one singular value that is equal to zero.
It is then trivial to show that $\b B^*\b B$ must have at least one eigenvalue that is equal to
zero, which means it is not an invertible matrix.

Another way of looking at the problem is that because the interferometer is not sensitive to part of
the southern hemisphere, there are infinitely many possible solutions to
Equation~\ref{eq:m-mode-matrix-equation} that will fit the measured data equally well.  We will
regularize the problem and apply an additional constraint that prefers a unique yet physically
reasonable solution.

\subsubsection{Tikhonov Regularization}

The process of Tikhonov regularization minimizes $\|\b v - \b B\b a\|^2 + \varepsilon\|\b a\|^2$ for
some arbitrary value of $\varepsilon > 0$ chosen by the observer. The solution that minimizes this
expression is given by
\begin{equation}\label{eq:tikhonov-solution}
    \atikh = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b v\,.
\end{equation}
Tikhonov regularization adds a small value $\varepsilon$ to the diagonal of $\b B^*\b B$, fixing the
matrix's singularity. By using the singular value decomposition (SVD) of the matrix $\b B = \b U \b
\Sigma \b V^*$. Equation~\ref{eq:tikhonov-solution} becomes
\begin{equation}
    \atikh = \b V (\b\Sigma^2 + \varepsilon \b I)^{-1}\b\Sigma \b U^*\b v\,,
\end{equation}
where
\[
    \b\Sigma = \left(
        \begin{array}{ccc}
            \sigma_1 & & \\
                     & \sigma_2 & \\
                     & & \ddots \\
        \end{array}
    \right)\,.
\]
The diagonal elements of $\b\Sigma$ are the singular values of $\b B$. The contribution of each
singular component to the Tikhonov-regularized solution is scaled by $\sigma_i / (\sigma_i^2 +
\varepsilon)$, where $\sigma_i$ is the singular value for the $i$th singular component. Tikhonov
regularization therefore acts to suppress any component for which
$\sigma_i\lesssim\sqrt{\varepsilon}$.  If $\sigma_i = 0$, the component is set to zero.

In practice the measurement $\b v$ is corrupted by noise with covariance $\b N$. For illustrative
purposes we will assume that $\b N=n\b I$ for some scalar $n$. In this case the covariance of the
Tikhonov-regularized spherical harmonic coefficients is
\begin{equation}
    \b C = n \b V (\b\Sigma^2 + \varepsilon\b I)^{-2} \b\Sigma^2 \b V^*\,.
\end{equation}
Each singular component is scaled by a factor of $\sigma_i^2/(\sigma_i^2 + \varepsilon)^2$.  In the
absence of Tikhonov regularization ($\varepsilon=0$) singular components with the smallest singular
values -- the ones that the interferometer is the least sensitive to -- actually come to dominate
the covariance of the measured spherical harmonic coefficients. Tikhonov regularization improves
this situation by downweighting these components.

While Tikhonov regularization will force unmeasured modes to zero, if a prior map of the sky already
exists, it would be preferrable to instead minimize $\|\b v - \b B\b a\|^2 + \varepsilon \|\b a-\b
a_\text{prior}\|^2$.

\subsubsection{L-Curves}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/lcurve/lcurve}
    \caption{
        An example L-curve computed from OVRO-LWA data at 36.528~MHz by trialing 200 different
        values of the regularization parameter $\varepsilon$. The $x$-axis is the norm of the
        solution (in this case the spherical harmonic coefficients) given in arbitrary units, and
        the $y$-axis is the least-squares norm given in arbitrary units Where the regularization
        parameter is small, the norm of the solution grows rapidly. Where the regularization
        parameter is large, the least-squares norm grows rapidly.
    }
    \label{fig:lcurve}
\end{figure}

Tikhonov regularization requires the observer to pick the value of $\varepsilon$. If $\varepsilon$
is too large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. However if $\varepsilon$ is too small then the problem will be
poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\varepsilon$ therefore requires understanding the trade-off between the two norms.

This trade-off can be analyzed quantitatively by trialing several values of $\varepsilon$, and
computing $\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in
Figure~\ref{fig:lcurve}. The shape of this curve has a characteristic L-shape, and as a result this
type of plot is called an L-curve. The ideal value of $\varepsilon$ lies near the turning point of
the plot. At this point a small decrease in $\varepsilon$ will lead to an undesired rapid increase
in $\|\b a\|^2$, and a small increase in $\varepsilon$ will lead to an undesired rapid increase in
$\|\b v - \b B\b a\|^2$.

In practice, the L-curve should be used as a guide to estimate a reasonable value of $\varepsilon$.
However better results can often be obtained by tuning the value of $\varepsilon$. For instance
increasing the value of $\varepsilon$ can improve the noise properties of the map by down-weighting
noisy modes. Decreasing the value of $\varepsilon$ can improve the resolution in the map by
up-weighting the contribution of longer baselines, which are likely fewer in number. In this respect
choosing the value of $\varepsilon$ is analagous to picking the weighting scheme in traditional
imaging where robust weighting schemes can be tuned to similar effect \citep{briggs}.  For the
OVRO-LWA we selected $\varepsilon = 0.01$ across all frequency channels. The distribution of
singular values of the transfer matrix with respect to $\sqrt{\varepsilon}$ is summarized in
Table~\ref{tab:summary}.

\subsubsection{The Moore-Penrose Pseudoinverse}

The Moore-Penrose pseudoinverse (denoted in this paper with a superscript $\dagger$), is commonly
applied to find the minimum-norm linear-least squares solution to a set of linear equations. This
can be used in place of Tikhonov regularization as
\begin{equation}
    \b{\hat a}_\text{Moore-Penrose} = \b B^\dagger\b v\,.
\end{equation}
Much like Tikhonov regularization, the Moore-Penrose pseudoinverse sets components with small
singular values (below some user-defined threshold) to zero. Components with large singular values
(above the user-defined threshold) are included in the calculation at their full amplitude with no
down-weighting of modes near the threshold. The essential difference between using the Moore-Penrose
pseudoinverse and Tikhonov regularization is that the pseudoinverse defines a hard transition from
on to off. Modes are either set to zero or included in the map at their full amplitude. On the other
hand Tikhonov regularization smoothly interpolates between these behaviors. Because of this,
Tikhonov regularization tends to produce better results in practical applications.

\subsubsection{Other Regularization Schemes}

In compressed-sensing applications, the $L^1$-norm is used in place of the $L^2$-norm in order to
encourage sparsity in the reconstructed signal. By expressing the sky in terms of spherical
harmonics there is no such expectation of sparsity in the solution, and so the $L^1$-norm is
inappropriate for this application.

Wiener filtering can also be used in place of Tikhonov regularization in the event that the observer
has a prior expectation on the covariance of the spherical harmonic coefficients in order to
interpolate missing information. This technique is demonstrated for simulated measurements by
\citet{2016arXiv161203255B}.

In this work we elect to pursue Tikhonv regularization due to its computational simplicity, smooth
down-weighting of noisy modes, and independence of prior information.

\subsection{CLEAN}\label{sec:clean}

In traditional radio astronomy imaging, CLEAN \citep{1974A&AS...15..417H} is a physically motivated
algorithm that interpolates between measured visibilities on the $uv$-plane. In the absence of this
interpolation, gaps in the interferometer's $uv$-coverage are assumed to be zero, and -- in the
image plane -- sources are convolved with a point spread function (PSF) that is characteristic of
the $uv$-coverage.  Fundamentally, the interferometer's PSF is determined by which modes were
assumed to be zero in the initial imaging process.

In $m$-mode analysis imaging, we assumed modes were zero in two separate ways:
\begin{enumerate}
    \item We selected a set of spherical harmonic coefficients $a_{lm}$ to describe the sky
        brightness distribution. All modes with $l>l_\text{max}$ are neglected and assumed to be
        zero.
    \item Tikhonov regularization forces linear combinations of spherical harmonic coefficients with
        $\sigma_i \lesssim \sqrt{\varepsilon}$ towards zero.
\end{enumerate}
As a consequence the final map of the sky is not assembled from a complete set of spherical
harmonics. Therefore just as in traditional imaging, $m$-mode analysis imaging produces dirty maps
in which sources are convolved with a PSF.  This PSF can be improved by increasing the number and
variety of baselines, which increases the number of modes for which $\sigma_i \gg
\sqrt{\varepsilon}$.  Alternatively by collecting more data, the signal-noise ratio of the measured
$m$-modes increases, which allows the observer to lower the value of $\varepsilon$ without
increasing the noise in the maps.  Finally, the CLEAN algorithm can be applied to interpolate some
of the missing information that was assumed to be zero.

The PSF of a dirty $m$-mode analysis map may be computed with
\begin{equation}\label{eq:psf}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the PSF at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta, \phi)$ is the
vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given by
\begin{equation}
    \b a_\text{PS}(\theta, \phi) = \begin{pmatrix}
        \vdots \\
        Y_{lm}^*(\theta, \phi) \\
        \vdots \\
    \end{pmatrix}
    = \begin{pmatrix}
        \vdots \\
        Y_{lm}^*(\theta, 0)\times e^{im\phi} \\
        \vdots \\
    \end{pmatrix} \,.
\end{equation}
In general the PSF can be a function of the right ascension and declination. However point sources
at the same declination take the same track through the sky and therefore (barring any ionospheric
effects) will have the same PSF. The PSF is therefore only a function of the declination. For
example, sources at low elevations will tend to have an extended PSF along the north-south axis due
to baseline foreshortening. For the OVRO-LWA antenna configuration
(Figure~\ref{fig:antenna-layout}), example PSFs at three separate frequencies are shown in
Figure~\ref{fig:psf}.  Adapting CLEAN for $m$-mode analysis requires either pre-computing
Equation~\ref{eq:psf} at a grid of declinations, or a method for rapidly evaluating
Equation~\ref{eq:psf} on the fly.

%\citep{2011MNRAS.413..411B} % I think Lincoln wants me to cite this paper

For an interferometer with more baselines than spherical harmonics used in the maps (eg. the
OVRO-LWA), $\b B^*\b B$ can be a much smaller matrix than the full transfer matrix $\b B$. Therefore
pre-computing $\b B^*\b B$ can allow the entire matrix to fit into memory on a single machine. This
greatly reduces the amount of disk I/O necessary for solving Equation~\ref{eq:psf}.

Additionally we can pre-compute the Cholesky decomposition of $\b B^*\b B + \varepsilon\b I = \b
U^*\b U$, where $\b U$ is an upper-triangular matrix. Inverting an upper triangular matrix is an
$\mathcal{O}(N^2)$ operation (instead of $\mathcal{O}(N^3)$ for a general matrix inverse).\footnote{
    Instead of computing $\b A^{-1}$, we solve the linear equation $\b A\b x = \b b$ each time the
    matrix inverse is needed so as to avoid numerical instabilities.
}
Equation~\ref{eq:psf} can then be rapidly evaluated from right to left as:
\begin{equation}\label{eq:rapid-psf}
    \b a_\text{PSF} =
        \b U^{-1}\,\big(\b U^*\big)^{-1}\,\big(\b B^*\b B\big)\,\b a_\text{PS}\,.
\end{equation}
Furthermore, Equation~\ref{eq:rapid-psf} does not need to be separately evaluated for each CLEAN
component. Instead we can identify $N$ CLEAN components, accumulate $\b a_\text{PS}$ for each
component, and evaluate Equation~\ref{eq:rapid-psf} on the accumulation. This can greatly reduce the
number of times this equation needs to be evaluated, but care must be taken to ensure that the $N$
components are not so close together that sidelobes from one may interact with another.

%\todo{describe blurring to avoid cleaning the diffuse emission}

Altogether the adaptation of CLEAN applied to the maps presented in this paper is summarized below.
\begin{algorithmic}[1]
    \Require{$\b a$ is the solution to Equation~\ref{eq:tikhonov-solution}}
    \Function{CLEAN}{$\b a$}
    \Let{$\b M$}{$\b B^*\b B$}
    \Let{$\b U$}{${\rm chol}(\b M + \varepsilon\b I)$} \Comment{Cholesky decomposition}
    \While{noise in map $>$ threshold}
    \State find $N$ pixels with the largest residual flux
    \Let{$\b x$}{$\sum_{i=1}^N \,(\text{pixel flux}) \times \b a_\text{PS}(\theta_i, \phi_i)$}
    \Let{$\b y$}{$\b U^{-1}\big(\b U^*\big)^{-1}\b M\b x$}
    \Let{$\b a$}{$\b a - (\text{loop gain})\times\b y$}
    \State record subtracted components
    \EndWhile
    \Let{$\b a$}{$\b a + (\text{restored components})$}
    \State \Return{$\b a$}
    \EndFunction
\end{algorithmic}

In summary, Tikhonov-regularized $m$-mode analysis imaging constructs a widefield synthesis image of
the sky from a complete earth rotation, and  with exact treatment of widefield-effects. This is
accomplished by solving a regularized block-diagonal matrix equation
(Equation~\ref{eq:tikhonov-solution}). The solution to this equation generates a map where
sources are convolved with a PSF characteristic of the interferometer (a function of the frequency,
antenna response, and baseline distribution with a full Earth rotation). The CLEAN algorithm is
adopted to deconvolve the PSF and produce the final sky maps.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations}\label{sec:observations}

\subsection{The Owens Valley Long Wavelength Array}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/antenna-layout/antenna-layout}
    \caption{
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 triangles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to $\sim1.5$ km. The red dots
        are core antennas that are disconnected from the correlator in order to accommodate these
        antennas. The 5 crosses are antennas equipped with noise-switched front ends.
    }
    \label{fig:antenna-layout}
\end{figure}

\begin{figure*}[t]
    \begin{tabular}{c}
        \includegraphics[width=\textwidth]{figures/psf/psf+75} \\
        \includegraphics[width=\textwidth]{figures/psf/psf+45} \\
        \includegraphics[width=\textwidth]{figures/psf/psf+0} \\
    \end{tabular}
    \caption{
        The $m$-mode analysis imaging PSF at three declinations (top row: $\delta=+75^\circ$, middle
        row: $\delta=+45^\circ$, bottom row: $\delta=+0^\circ$) and three frequencies (left column:
        36.528~MHz, middle column: 52.224~MHz, right column: 73.152~MHz).  The PSF is computed by
        evaluating Equation~\ref{eq:psf}. Above 55~MHz, the angular extent of the PSF does not
        follow the expected scaling with frequency because the angular resolution is limited by the
        selection of $l_\text{max}=1000$. The FWHM at $\delta=+45^\circ$ is listed in
        Table~\ref{tab:summary}.
    }
    \label{fig:psf}
\end{figure*}

The Owens Valley Long Wavelength Array (OVRO-LWA) is a 288-element interferometer located at the
Owens Valley Radio Observatory near Big Pine, California \citep{hallinan_2017}.  The OVRO-LWA is a
low-frequency instrument with instantaneous bandwidth covering 27~MHz to 85~MHz and with 24~kHz
channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that there are
$288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA correlator
\citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible hemisphere in a
single snapshot image.

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging (see Figure~\ref{fig:antenna-layout}).  Of these 288 antennas, 251 are contained
within a 200 m diameter core, 32 are placed outside of the core in order to extend the maximum
baseline length to $\sim$1.5~km, and the final 5 antennas are equipped with noise-switched front
ends for calibrated total power measurements of the global sky brightness.  These antennas are used
as part of the LEDA experiment \citep{2017arXiv170909313P} to measure the global signal of 21~cm
absorption from the cosmic dawn. In the current configuration 32 antennas (64 signal paths) from the
core are disconnected from the correlator in order to accommodate the 32 antennas on longer
baselines. A final stage of construction will involve 64 additional antennas installed on long
baselines out to a maximum length of 2.6~km.

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
During this time the OVRO-LWA operated as a zenith-pointing drift scanning interferometer.  The
correlator dump time was selected to be 13 seconds such that the correlator output evenly divides a
sidereal day. Due to the computational considerations presented in \S\ref{sec:mmode-analysis}, eight
24~kHz channels are selected for imaging from this dataset: 36.528~MHz, 41.760~MHz, 46.992~MHz,
52.224~MHz, 57.456~MHz, 62.688~MHz, 67.920~MHz, and 73.152~MHz.

\subsection{Complex Gain Calibration}\label{sec:gaincal}

The complex gain calibration is responsible for correcting per-antenna amplitude and phase errors.
This is accomplished using a sky model and a variant of alternating least squares colloquially known
as ``Stefcal''
\citep{2008ISTSP...2..707M, 2014A&A...571A..97S}\footnote{
    The calibration routine is written in the Julia programming language
    \citep{doi:10.1137/141000671}, and is publicly available online
    (\url{https://github.com/mweastwood/TTCal.jl}) under an open source license (GPLv3 or any later
    version).
}.

Cyg~A and Cas~A are -- by an order of magnitude -- the brightest point-like radio sources in the
northern hemisphere. Therefore the optimal time to solve for the interferometer's gain calibration
is when these sources are at high elevations.  The antenna complex gains are measured from a 22
minute track of data when Cyg~A and Cas are at high elevations. The gains measured in this way are
then used to calibrate the entire 28 hour dataset. The calibration sky model consists only of Cyg~A
and Cas~A. The model flux of Cyg~A is set to the \citet{1977A&A....61...99B} spectrum while the flux
of Cas~A is measured from the data itself (using a preliminary calibration solved for with a
fiducial Cas~A spectrum).

Calibrating in this manner generates $\sim$ arcminute errors in the astrometry of the final sky maps
due to ionospheric refractive offsets during the time of calibration.  These residual errors in the
astrometry are corrected post-imaging by registering the images with respect to all VLSSr
\citep{2014MNRAS.440..327L} sources that are bright ($>30$~Jy with a consistent flux density
measured with the OVRO-LWA), and not too close to other bright sources (at least $1^\circ$
separation).

Temperature fluctuations of the analog electronics generate 0.1~dB sawtooth oscillations in the
analog gain. These oscillations occur with a variable 15 to 17 minute period associated with HVAC
cooling cycles within the electronics shelter that houses these electronics.  The amplitude of these
gain fluctuations is calibrated by smoothing the autocorrelation amplitudes on 45 minute timescales.
The ratio of the measured auto-correlation power to the smoothed auto-correlation power defines a
per-antenna amplitude correction that is then applied to the cross-correlations.

\subsection{Primary Beam Measurements}\label{sec:beam}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/beam/beam}
    \caption{
        Empirical fits to the OVRO-LWA Stokes-I primary beam (the response of the $x$- and
        $y$-dipoles has been summed) at three frequencies: 36.528 MHz (left panel), 52.224 MHz
        (middle panel), and 73.152 MHz (right panel). The source tracks used to measure the beam
        model are overlaid. From north to south, these tracks correspond to Cas~A, Cyg~A, Per~B,
        Tau~A, Vir~A, Her~A, 3C~353, and Hya~A.  The fitting process is described in
        \S\ref{sec:beam}, and residuals for Cyg~A and Cas~A are in Figure~\ref{fig:scintillation}.
    }
    \label{fig:beam}
\end{figure*}

In order to generate wide-field images of the sky, the response of the antenna to the sky must be
known. Drift-scanning interferometers like the OVRO-LWA can empirically measure their primary beam
under a mild set of symmetry assumptions \citep{2012AJ....143...53P}. The symmetry assumptions are
necessary to break the degeneracy between source flux and beam amplitude when the flux of the source
is unknown. In this work we assume symmetries that are apparent in the antenna design, but
real-world defects and coupling with nearby antennas will contribute towards breaking these
symmetries at some level. In particular we assume that the $x$- and $y$-dipoles have the same
response to the sky after rotating one by 90$^\circ$, and that the beam is invariant under
north-south and east-west flips.

We measure the flux of several bright sources (Cyg~A, Cas~A, Tau~A, Vir~A, Her~A, Hya~A,
Per~B\footnote{
    At 74~MHz, Per~B is about two times brighter than Per~A.
}, and
3C~353) as they pass through the sky and then fit a beam model composed of Zernike polynomials to
those flux measurements. We select the basis functions to have the desired symmetry ($Z_0^0$,
$Z_2^0$, $Z_4^0$, $Z_4^4$, $Z_6^0$, $Z_6^4$, $Z_8^0$, $Z_8^4$, and $Z_8^8$) and the beam amplitude
at zenith is constrained to be unity. See Figure~\ref{fig:beam} for an illustration of a fitted beam
model at several frequencies. This process is repeated for each frequency channel.

\subsection{Ionospheric Conditions}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/scintillation-refraction/scintillation-refraction}
    \caption{
        (a, b) These panels show the measured apparent flux of Cyg~A and Cas~A at 36.528 MHz (red
        points) and 73.152 MHz (blue points) as a function of time over the observing period. The
        solid black curves show the expected flux computed using the empirical beam model fits. The
        thermal noise contribution to each point is about 50 Jy.  Cyg~A is occulted by the White
        Mountains when it is low on the horizon to the east.
        (c, d) These panels show the measured position offset of Cyg~A and Cas~A relative to their
        true astronomical positions at 36.528 MHz (red line) and 73.152 MHz (blue line).
    }
    \label{fig:scintillation}
\end{figure*}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/vtec/vtec}
    \caption{
        The median vertical TEC within 200 km of OVRO during the time of the observation. The gray
        shaded regions indicate times outside of the observing period. The gray vertical lines
        indicate sunrise and sunset (as labeled).
    }
    \label{fig:vtec}
\end{figure}

The geomagnetic conditions during this time were mild. The Disturbance storm time (Dst) index, which
measures the $z$-component of the interplanetary magnetic field, was
$>-30$ nT during the entirety of the observing period.\footnote{
    The Dst index was obtained from the World Data Center for Geomagnetism, Kyoto University
    (\url{http://swdcwww.kugi.kyoto-u.ac.jp/}). Accessed 2017-07-25.
}
Following the classification scheme of \citet{2008GMS...181.....K}, a weak geomagnetic storm has
$\text{Dst} < -30$ nT. Stronger geomagnetic storms have $\text{Dst} < -50$ nT.

Despite the mild conditions, low frequency interferometric observations are still affected by the
index of refraction in the ionosphere.  Figure~\ref{fig:vtec} shows the median vertical total
electron content (TEC) above OVRO measured from GPS \citep{1999JASTP..61.1205I}. The median is
computed over all GPS measurements within 200 km of the observatory. Over the observing period the
TEC smoothly varies from 20 TECU at midday to 5 TECU during the night. However, this measurement is
only sensitive to large scale fluctuations in the ionosphere and does not capture small scale
fluctuations.

Small scale fluctuations are best characterized by source scintillation and refractive offsets.
Figure~\ref{fig:scintillation} shows the apparent flux and position offset of Cyg~A and Cas~A as a
function of time over the entire observing period. Both sources exhibit rapid scintillation on the
timescale of a single integration (13 seconds). For example at 36.528 MHz, it is not unusual for Cyg
A to have a measured flux variations of 50\% between adjacent 13 second integrations.  The variance
at 36.528~MHz compared with the variance at 73.152~MHz is consistent with an ionospheric $\nu^{-2}$
origin. The measured position offset of each source is a measurement of the ionospheric phase
gradient across the array.  This varies on slower 10 minute timescales refracting each source by as
much as 20 arcmin (at 36.528 MHz) from their true astronomical positions as waves in the ionosphere
pass through the line of sight. At 74~MHz on the VLA, \citet{2007ApJS..172..686K} observe $\sim
1$~arcmin refractive offsets during the night, and $\sim 4$~arcmin offsets during the day on similar
$\sim10$ minute timescales.  The impact of these effects on the sky maps is simulated in
\S\ref{sec:ionosphere}.

\subsection{Source Removal}\label{sec:source-removal}

\subsubsection{Cygnus A and Cassiopeia A}

Due to the rapid and large ionospheric fluctuations seen in Figure~\ref{fig:scintillation}, CLEAN
cannot be relied on to accurately deconvolve bright sources.  However, without removing bright
sources from the data, sidelobes from these sources will dominate the variance in the sky maps.  At
74~MHz Cyg~A is a 15,000~Jy source \citep{1977A&A....61...99B}. A conservative estimate for the
confusion limit at 74 MHz with a 15~arcmin beam is 1000~mJy \citep{2014MNRAS.440..327L}. Therefore
we require that Cyg~A's sidelobes be at least $-45$~dB down from the main lobe to prevent Cyg~A's
sidelobes from dominating the variance in the image.

In order to account for propagation effects through the ionosphere direction-dependent calibration
and peeling \citep{2008ISTSP...2..707M, 2015MNRAS.449.2668S} must be used.  Direction-dependent
calibration allows the per-antenna amplitude and phase towards a bright astronomical source to be
free parameters. In the dataset used in this paper, scintillation and refractive-offset events on
the timescale of a single integration (13 seconds) are observed (Figure~\ref{fig:scintillation}).
Therefore the direction dependent calibration changes on these timescales and we must solve for one
set of complex gains per source per integration.

The largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular scale of Cyg~A is
$\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines at 73~MHz, the OVRO-LWA
marginally resolves both sources. A resolved source model is needed for both sources. We fit a
self-consistent resolved source model to each source. This is performed by minimizing the variance
within an aperture located on each source after peeling. By phasing up a large number of
integrations before imaging (at least 1 hour) it is possible to smear out the contribution of the
rest of the sky.  We then use NLopt's Sbplx routine \citep{nlopt, sbplx} to vary the parameters in a
source model until the variance within the aperture is minimized. Cyg~A is modeled with 2 Gaussian
components, while Cas~A is modeled with 3~Gaussian components. Armed with these  source models, both
sources can be peeled from the dataset to the required dynamic range.

\subsubsection{Other Bright Sources}

Other bright sources namely Vir~A, Tau~A, Her~A, Hya~A, Per~B, and 3C~353 are also removed from
the visibilities prior to imaging. Because these sources are much fainter than Cyg~A and Cas~A, we
do not need resolved source models to be able to remove these sources from the visibilities without
residual sidelobes contaminating the image.

However, the ionosphere will cause these sources to scintillate and refract. The position and flux
of each source is measured separately in each channel and integration. The sources are then
subtracted from the visibilities using the updated position and flux of the source. The brightest of
these sources (Vir~A and Tau~A) are peeled using a direction dependent calibration when they are at
high elevations.

\subsubsection{The Sun}

The Sun can be trivially removed from any map of the sky by constructing the map using only data
collected at night. A map of the entire sky can be obtained by using observations spaced 6 months
apart.  However the dataset used in this paper consists of 28 consecutive hours. Therefore we
attempt to suppress the contribution of the Sun to the maps with the goal of suppressing its
sidelobes.  The Sun is well-resolved by the OVRO-LWA and hence a detailed source model is needed. In
fact the optical depth $\tau=1$ surface of the Sun changes with frequency and as a consequence a new
model is needed at each frequency. While we could fit a limited number of Gaussian components to
Cyg~A and Cas~A, this is insufficient for the Sun.  Additionally while most astronomical sources at
these frequencies have negative spectral indices, the Sun has a positive spectral index. Therefore
more care will need to be taken in subtracting the Sun at higher frequencies than at lower
frequencies.

The strategy used for removing the Sun below 55 MHz involves fitting a shapelet
\citep{2003MNRAS.338...35R} model to the Sun and subtracting without the use of direction dependent
gains. The shapelet fitting is performed in the visibility space. Above 55 MHz a model is fit to the
Sun by minimizing the residuals after peeling (in the same way that models are obtained for Cyg~A
and Cas~A). The Sun is then peeled from each integration using direction dependent gains.

\subsection{Flux Scale}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/flux-scale/flux-scale}
    \caption{
        Measured fluxes (black points) of 11 sources plotted against the published spectra from
        \citet{2017ApJS..230....7P} (solid line above 50 MHz, dotted line below 50 MHz),
        \citet{2012MNRAS.423L..30S} (dashed line), and \citet{1977A&A....61...99B} (dot-dash line).
    }
    \label{fig:flux-scale}
\end{figure*}

The flux scale of the data was tied to the \citet{1977A&A....61...99B} spectrum of Cyg~A during gain
calibration. However gain calibration is also a function of the beam model and the spectrum used for
Cas~A. Recent work by \citet{2012MNRAS.423L..30S} (hereafter SH12) using archival date from the
literature and \citet{2017ApJS..230....7P} (hereafter PB17) using the VLA has expanded the number of
low-frequency radio sources with calibrated flux measurements from one (just Cyg~A) to eleven in
total. While the SH12 flux scale is valid between 30 MHz and 300 MHz, the PB17 flux scale is
somewhat more limited because the lowest frequency observations come from the VLA 4-band system. As
a consequence the PB17 flux scale is not valid below 50 MHz.

Figure~\ref{fig:flux-scale} shows a comparison between flux measurements made using the all-sky maps
from this work, and spectra from the aforementioned flux scales. Generally the OVRO-LWA flux
measurements agree to between 5 and 10\% with the SH12 spectra. Below 50 MHz there can be
substantial departures with respect to the extrapolated PB17 spectra (for eg.  3C 286, 3C 295, and
3C 380), but it is usually the case that we have much better agreement with the SH12 spectra. This
indicates that the PB17 spectra cannot be extrapolated below 50 MHz.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:results}

\begin{figure*}[ht]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw04} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw06} \\
    \end{tabular}
    \caption{
        These 8 panels illustrate (with a Mollweide projection and logarithmic color scale) the 8
        full-sky maps generated with Tikhonov-regularized $m$-mode analysis imaging and the
        OVRO-LWA.  Each map covers the sky north of $\delta=-30^\circ$ with angular resolution of
        $\sim15$ arcmin. Eight bright sources have been removed from each map (Cyg~A, Cas~A, Vir~A,
        Tau~A, Her~A, Hya~A, Per~B, and 3C~353). The small blank region near $l=+45.7^\circ$,
        $b=-47.9^\circ$ corresponds to the location of the Sun during the observation period.  A
        detailed summary of the properties of each map is given in Table~\ref{tab:summary}.
    }
    \label{fig:channel-maps}
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw08} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw10} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw12} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw14} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw16} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw18} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/spectral-index/ovro-lwa-sky-map.pdf}
    \caption{
        This Mollweide-projected map is constructed from 3 maps of the sky at 36.528~MHz (red),
        52.224~MHz (green), and 73.152~MHz (blue). The maps are scaled by $\nu^{2.5}$ before
        combining and the color scale is logarithmic (as in Figure~\ref{fig:channel-maps}).
        Therefore regions with a spectral index of $-2.5$ appear white. Areas with a steeper index
        appear red and areas with a flatter spectral index appear blue.
    }
    \label{fig:three-color}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/spectral-index/better-internal-spectral-index}
    \caption{
        \todo{updated this caption}
        This plot shows the local spectral index (estimated by means of a local T-T plot) as a
        contour plot. The contours are overlaid on the three-color image from
        Figure~\ref{fig:three-color}. The spectral index fit along the galactic plane is masked
        because T-T plots assume a single emission mechanism operates in a given region. Along the
        galactic plane free-free dominated regions lie in close proximity to synchrotron dominated
        regions. This introduces large systematic errors to the power-law fit unless care is taken
        to cleanly separate the two types of emission mechanisms.
    }
    \label{fig:internal-spectral-index}
\end{figure*}

\begin{table*}[t]
    \centering
    \begin{tabular}{ccccccccc}
        \hline
        \hline
        & \tbf{Frequency} & \tbf{Bandwidth} & \tbf{FWHM}\footnote{
                The full-width half-maximum (FWHM) of the synthesized beam at $\delta=+45^\circ$.
            }&
            \multicolumn2c{\tbf{Noise}\footnote{
                Measured with a jackknife and splitting the dataset into even- and odd-numbered
                integrations. This estimate therefore includes all noise sources that act on the
                timescale of a single 13 second integration (eg. thermal, ionospheric, etc.).
            }} &&
            \multicolumn2c{\tbf{Fraction of Modes}\footnote{
                Singular values of the transfer matrix compared with the value of the regularization
                parameter $\varepsilon$ used while solving Equation~\ref{eq:tikhonov-solution}. As
                discussed in the text, singular vectors with corresponding singular values $\sigma
                \ll \sqrt{\varepsilon}$ are set to zero by the Tikhonov regularization procedure.
            }} \\
        \cline{5-6}
        \cline{8-9}
        \tbf{\#} & MHz & MHz & arcmin &
            K & mJy/beam &&
            with $\sigma>\sqrt{\varepsilon}$ &
            with $\sigma>\sqrt{10\varepsilon}$ \\
        \hline
        1 & 36.528 & 0.024 & 18.5 & 595. & 799. && 0.391 & 0.077 \\
        2 & 41.760 & 0.024 & 17.2 & 541. & 824. && 0.480 & 0.070 \\
        3 & 46.992 & 0.024 & 16.3 & 417. & 717. && 0.504 & 0.056 \\
        4 & 52.224 & 0.024 & 15.6 & 418. & 814. && 0.535 & 0.053 \\
        5 & 57.456 & 0.024 & 15.4 & 354. & 819. && 0.542 & 0.049 \\
        6 & 62.688 & 0.024 & 15.3 & 309. & 843. && 0.540 & 0.046 \\
        7 & 67.920 & 0.024 & 15.3 & 281. & 894. && 0.529 & 0.047 \\
        8 & 73.152 & 0.024 & 15.7 & 154. & 598. && 0.512 & 0.051 \\
        \hline \hline
    \end{tabular}
    \caption{A summary of the generated all-sky maps}
    \label{tab:summary}
\end{table*}

Eight sky maps were constructed using Tikhonov-regularized $m$-mode analysis imaging and cleaning
with observations from the OVRO-LWA. Each map is individually shown in
Figure~\ref{fig:channel-maps}, and Figure~\ref{fig:three-color} is a three-color image constructed
from the maps at 36.528~MHz, 52.224~MHz, and 73.152~MHz. The maps cover the sky north of
$\delta=-30^\circ$ with $\sim 15$ arcmin angular resolution resolution. The 8 brightest northern
hemisphere point sources are removed from each map (Cyg~A, Cas~A, Vir~A, Tau~A, Her~A, Hya~A, Per~B,
and 3C~353) as described in \S\ref{sec:source-removal}, and there is a small blank region near
$l=+45.7^\circ$, $b=-47.9^\circ$ corresponding to the position of the Sun during the observing
window. The properties of each map, including frequency, bandwidth, angular resolution, and thermal
noise are presented in Table~\ref{tab:summary}.

Due to the considerations presented by \citet{2016ApJ...826..116V}, each of these maps is monopole
subtracted ($a_{00}=0$). Consequently a simple pixel-by-pixel powerlaw fit will be unduly affected
by the choice of zero-level correction. Instead, temperature-temperature plots (T-T plot) can be
used to measure the spectral index independently of any zero-level corrections
\citep{1962MNRAS.124..297T}. This method relies on the assumption that all pixels in a given region
are described by the same power law. In Figure~\ref{fig:internal-spectral-index} the spectral index
is locally estimated within a region $\sim10^\circ$ across by constructing a local T-T plot at every
pixel between the maps at 36.528~MHz and 73.152~MHz. The spectral index contours are overlaid on the
3-color image from Figure~\ref{fig:three-color} where free-free regions responsible for flattening
the spectral index in the galactic plane are plainly visible as extended blue blobs.  Similarly the
spectral index can be seen to steepen as galactic latitude increases (this is a transition from blue
to white to yellow that is also evident in the spectral index contours).

In order to suppress sources of terrestrial interference, all spherical harmonics with $m=0$, or
$m=1$ and $l>100$ are filtered from the map (where the spherical harmonics are defined in the J2017
coordinate system). As will be discussed in \S\ref{sec:rfi}, these spherical harmonics are
particularly susceptible to contamination by RFI and common-mode pickup. As a consequence,
astronomical emission that circles the J2017 north celestial pole (NCP) is filtered from the maps.
This filtering creates negative rings around the NCP at the declination of bright point sources.
These rings are naturally removed from the map during CLEANing as long as this filtering step is
including in the PSF calculation.

The noise in each map is empirically measured using jackknife resampling. The dataset is split into
even- and odd-numbered integrations. These two groups are imaged and cleaned independently, before
being compared against the maps constructed from all of the available data using the standard
jackknife variance estimator. This estimate of the variance includes all sources of error that
operate on $\sim13$ second timescales (the integration time) such as thermal noise in rapid
ionospheric fluctuations, but does not account for more slowly varying effects (for example,
sidereal variation in the system temperature or day-night fluctuations in the ionosphere). These
noise calculations are summarized in Table~\ref{tab:summary}.  VLSSr source counts
\citep{2014MNRAS.440..327L} suggest that the confusion limit at 74 MHz and 15 arcmin angular
resolution is $\sim 1000\times(\nu/74\,{\rm MHz})^{-0.7}$ mJy.  Each channel map achieves thermal
noise $<900$ mJy and therefore each map is likely at or near the confusion limit.

Each map from Figure~\ref{fig:channel-maps} is freely available online in Healpix format
\citep{2005ApJ...622..759G} on LAMBDA (\url{https://lambda.gsfc.nasa.gov/}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Error Analysis}\label{sec:error}

\subsection{Terrestrial Interference and Pickup}\label{sec:rfi}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/terrestrial-interference/smeared}
    \caption{
        Terrestrial sources of correlated noise that are apparent after averaging the visibilities
        at 62.688 MHz over the entire 28 hour observing period (keeping the phase center at zenith
        such that astronomical sources of radio emission are smeared along tracks of constant
        declination). Each panel represents a different component that is removed from the
        visibilities. The images are generated using WSClean \citep{2014MNRAS.444..606O}, uniform
        weighting, and only baselines longer than 15 wavelengths. Panels (a) and (b) illustrate
        components that appear noise-like in image-space, but are in fact a constant-offset to the
        measured visibilities likely associated with cross-talk or common-mode pickup. Panel (c)
        illustrates a component that is clearly associated with an RFI source on the horizon to the
        west-north-west of the OVRO-LWA. This RFI source is likely an arcing power line.
        Figure~\ref{fig:rings} illustrates the characteristic ring-like artifacts introduced into
        the maps if these 3 components are not removed prior to $m$-mode analysis imaging. The
        component shown in panel (a) has about twice the amplitude ($\|\b v_\text{terrestrial}\|$)
        of panels (b) and (c), and for all three components $\|\b B^*\b v_\text{terrestrial}\|/(\|\b
        B\|\|\b v_\text{terrestrial}\|) \sim 0.035$.
    }
    \label{fig:fitrfi}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/terrestrial-interference/rings}
    \caption{
        This is a Mollweide-projected image of the artifacts introduced to the $m$-mode analysis
        maps by the three terrestrial sources shown in Figure~\ref{fig:fitrfi}. Because these
        sources are not moving through the sky sidereally, they tend to be smeared along rings of
        constant declination. The spurs seemingly radiating from the north celestial pole are a
        Moir\'{e} pattern (ie. an artifact of the pixelization).
    }
    \label{fig:rings}
\end{figure*}

When writing Equation~\ref{eq:basic-imaging}, it is implicitly assumed that the correlated voltage
fluctuations measured between pairs of antennas are exclusively generated by astronomical sources of
radio emission. In practice this assumption can be violated. For instance a low-frequency
interferometer located in the vicinity of an arcing power line will see an additional contribution
from the radio-frequency interference (RFI) generated by the arcing process. Similarly common-mode
pickup along the analog signal path of the interferometer may generate an additional spurious
contribution to the measured visibilities. While the amplitude and phase of these contaminating
signals may fluctuate with time, they do not sweep across the sky at the sidereal rate
characteristic of astronomical sources.

The Owens Valley is an important source of water and power for the city of Los Angeles.
Unfortunately this means that high voltage power lines run along the valley $\gtrsim10$~km to the
west of the OVRO-LWA. Some of these power line poles have faulty insulators that arc and produce
pulsed, broadband RFI. Because these poles exist in the near-field of the array, we have been able
to localize some of them by using the curvature of the incoming wavefront to infer a distance.
Efforts are currently underway to work with the utility pole owners to have these insulators
replaced.

In the meantime it is possible to suppress their contamination in the dataset. The contribution of
these RFI sources to the visibilities can be plainly seen by averaging $>24$ hours of data with the
phase center set to zenith. In this way, true sky components are smeared along tracks of constant
declination while terrestrial sources (ie. the arcing power lines or any contribution due to
common-mode pickup) are not smeared.  Obtaining a model for the RFI is complicated by the fact that
the contaminating sources are at extremely low elevations where the antenna response is essentially
unknown (and inhomogeneous due to antenna shadowing effects). Therefore it is not enough to know the
physical location of the faulty insulator generating the RFI. In addition we must know the response
of each antenna (amplitude and phase) in the appropriate direction. This motivates the use of
peeling, which allows the antenna response to be a free parameter.  Therefore model visibilities for
the RFI can be obtained by peeling the sources after smearing the visibilities over $>24$ hours.
Figure~\ref{fig:fitrfi} shows an illustration of some of the removed components at 62.688 MHz.

While attempting to peel RFI sources from the averaged visibilities, it was discovered that
frequently peeling would remove components from the visibilities that are not obviously associated
with any source on the horizon or elsewhere in the sky (see panels (a) and (b) in
Figure~\ref{fig:fitrfi}).  These components appear noise-like in the images, but they are actually a
constant offset to the measured visibilities and are therefore likely associated with cross-talk or
some form of common-mode pickup. If these components are not subtracted from the measured
visibilities, they contribute ring-like structures into the sky maps as seen in
Figure~\ref{fig:rings}. This figure is not a simulation, but rather a difference between maps
created before and after measuring and subtracting the components in Figure~\ref{fig:fitrfi} from
each integration.

The first step in Equation~\ref{eq:tikhonov-solution} is to compute $\b B^*\b v$. In this step we
compute the projection of the measurement $\b v$ onto the space spanned by the columns of $\b B$.
Each column of $\b B$ describes the interferometer's response to a corresponding spherical harmonic
coefficient of the sky brightness distribution. Therefore the act of computing $\b B^*\b v$ is to
project the measured $m$-modes onto the space of $m$-modes which could be generated by astronomical
sources. The degree to which a source of terrestrial interferer will contaminate a map generated
using $m$-mode analysis imaging is determined by its amplitude after projection.

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = \b 0$.
In practice this is unlikely. In general the contamination is proportional to the overall amplitude
of the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/(\|\b B\|\|\b v_\text{terrestrial}\|)$).

These terrestrial sources do not rotate with the sky and hence their contamination tends to be
restricted to modes with small $m$. In this dataset the contamination is largely restricted to $m
\lesssim 1$. Although the RFI is capable of fluctuating on short timescales, in this case the
artifacts it introduces seem to be restricted to small $m$ (presumably because the phase is not
fluctuating).  As a result if the contamination is not suppressed, it will manifest itself as rings
along stripes of constant declination. This effect is plainly visible in Figure~\ref{fig:fitrfi}.
Because of the distinctive ring-like pattern created by terrestrial sources, we additionally chose
to discard spherical harmonics with either $m=0$ or $m=1$ and $l>100$ in order to further suppress
the contamination.

\subsection{The Ionosphere}\label{sec:ionosphere}

\begin{figure*}[t]
    \centering
    \begin{tabular}{cc}
        \includegraphics[width=\columnwidth]{figures/ionospheric-simulations/scintillation-4} &
        \includegraphics[width=\columnwidth]{figures/ionospheric-simulations/scintillation-18} \\
        \includegraphics[width=\columnwidth]{figures/ionospheric-simulations/refraction-4} &
        \includegraphics[width=\columnwidth]{figures/ionospheric-simulations/refraction-18} \\
    \end{tabular}
    %\includegraphics[width=\textwidth]{figures/ionospheric-simulations/ionospheric-simulations}
    \caption{
        This figure illustrates the corrupting influence of the ionosphere at 36.528~MHz (left
        column) compared with 73.152~MHz (right column). Each panel shows the \textit{simulated} PSF
        for a source at the location of Cas~A.  Each panel illustrates the percent difference
        (relative to the peak flux of the uncorrupted PSF) due to including an ionospheric effect.
        In the top row, the simulated source scintillates using the measured light-curve for Cas~A
        in Figure~\ref{fig:scintillation}. In the bottom row, the simulated source is refracted from
        its true position using the measured refractive offsets for Cas~A in
        Figure~\ref{fig:scintillation}.
    }
    \label{fig:ionospheric-simulations}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/postage-stamps/3C134}
    \caption{
        A zoom-in of 3C~134 at 36.528~MHz (left panel), 52.224~MHz (middle panel), and 73.152~MHz
        (right panel). At 36.528~MHz there are $\sim10\%$ artifacts around 3C~134 that persist after
        CLEANing due to ionospheric effects. As expected for an ionospheric origin, these artifacts
        decrease in amplitude as frequency increases.
    }
    \label{fig:3C134}
\end{figure*}

One of the key assumptions made by $m$-mode analysis is that the sky is static.  We assume that the
only time-dependent behavior is the rotation of the Earth, which slowly rotates the sky through the
fringe patterns of the interferometer. At low frequencies the ionosphere violates this assumption.
In particular, ionospheric scintillation and refractive offsets will cause even static sources to
exhibit significant variability (Figure~\ref{fig:scintillation}).

The correlation observed on a given baseline for a single point source is
\begin{equation}
    V_\nu(t_{\textrm{sidereal}}) = I_\nu B_\nu(t_{\textrm{sidereal}}),
\end{equation}
where $I_\nu$ is the flux of the source at the frequency $\nu$, and $B_\nu$ is the baseline transfer
function defined by Equation~\ref{eq:baseline-transfer-function}. The transfer function is a
function of the direction to the source, which is in turn a function of the sidereal time
$t_{\textrm{sidereal}}$. If the source is varying, from intrinsic variability or due to
scintillation, than the source flux is also a function of the time coordinate $t$ such that
\begin{equation}
    V_\nu(t_{\textrm{sidereal}}) = I_\nu(t) B_\nu(t_{\textrm{sidereal}}),
\end{equation}
where $t_{\textrm{sidereal}} = (t \mod 23^h56^m)$.

In order to compute the $m$-modes we must take the Fourier transform with respect to the sidereal
time. As a consequence of the Fourier convolution theorem we find:
\begin{equation}\label{eq:no-longer-block-diagonal}
    V_{\nu, m} \sim \sum_{m^\prime} V_{m^\prime}^\textrm{static} I_{\nu, m-m^\prime}\,.
\end{equation}
where $V_{\nu, m}^{\textrm{static}}$ is the set of observed $m$-modes if the source was actually
static, and $I_{\nu, m-m^\prime}$ is the Fourier transform of the light curve $I_{\nu}(t)$.
Equation~\ref{eq:no-longer-block-diagonal} indicates that power is scattered between different
values of $m$. As a consequence, the true transfer matrix, which is exactly block diagonal in the
ideal case, is no longer truly block diagonal \citep{richard_ionosphere_thoughts}.

The maps presented in Figure~\ref{fig:channel-maps} do not account for any off-diagonal terms
arising from ionospheric fluctuations. The effect of this can be seen in
Figure~\ref{fig:ionospheric-simulations}. In this simulation at 36.528~MHz, a point source is placed
at the location of Cas A. In the first panel the source flux traces the beam model exactly and the
source is always located exactly at the location of Cas A. In the middle panel, the source is
allowed to scintillate in the same way Cas A does in Figure~\ref{fig:scintillation}. The image of
the point source is effectively unchanged from the ideal PSF. In the right panel, the source
position is allowed to vary in the same way Cas A does in Figure~\ref{fig:scintillation}. The
refractive offsets lead to a clear degradation of the sidelobes. Because the sidelobes are altered
from that of the ideal PSF, refractive offsets will restrict the dynamic range it is possible to
obtain with the CLEAN algorithm described in \S\ref{sec:clean}. This effect can be clearly seen in
Figure~\ref{fig:3c134}, where 10\% errors within 1$^\circ$ of 3C~134 are seen at 36.528~MHz.  As
expected for an ionospheric effect these errors decrease to a few percent at 52.224~MHz, and less at
73.152~MHz. We therefore conclude that ionospheric effects directly limit the dynamic range in the
vicinity of bright point sources.

\subsection{Beam Errors}

A model of the antenna beam is essential for widefield imaging. Because $m$-mode analysis imaging
operates on a full sidereal day of data, images are constructed after watching each point in the sky
move through a large slice of the beam (excepting the celestial poles). The beam model therefore
serves two purposes:

\begin{enumerate}
    \item setting the flux scale as a function of declination
    \item reconciling observations from two separate sidereal times
\end{enumerate}

In the first case, all sources at a given declination take the same path through the antenna primary
beam. If the antenna response is overestimated along this track then all sources at this declination
will have their flux underestimated. Similarly if the antenna response is underestimated then all
the sources will have their flux overestimated. The errors in Figure~\ref{fig:flux-scale} do not
show a clear pattern with declination. Two sources have a clear systematic offset at all
frequencies: 3C 353 and 3C 380. 3C 353 is the second southernmost source, but Hya A -- the first
southernmost source -- does not exhibit this systematic error. Similarly 3C 380 is at a comparable
declination to Lyn A, which appears, if anything, to have its flux overestimated. The absence of a
coherent pattern does not eliminate the possibility of beam errors effecting the flux scale, but it
does mean that these errors are at least comparable to the errors inherent to the flux scale itself.

The second case is more subtle. Sources are observed at a wide range of locations in the primary
beam of the antenna. The imaging process must reconcile all of these observations together, and the
beam model provides the instructions for how to do this. In the event of an error in the beam model,
it can be expected that the beam will introduce errors into the sky maps that will limit the dynamic
range in the vicinity of bright point sources.  \citet{2015PhRvD..91h3514S} simulate the effect of
beam errors on a cosmological analysis concluding that the beam must be known to one part in $10^4$.
Our requirements are significantly less stringent because we are estimating the sky brightness
instead of estimating the amplitude of a faint cosmological signal in the presence of foreground
emission that dominates the signal by 5 orders of magnitude. In fact in \S\ref{sec:ionopshere} we
found that ionospheric effects likely dominate over other sources of error that affect the PSF
shape. Therefore we conclude that the beam models generated in \S\ref{sec:beam} are sufficient to
limit the effect of beam errors on the PSF to at least less than those introduced by the ionosphere.

\subsection{Polarization Leakage}

\citet{2015PhRvD..91h3514S} describe how to generalize $m$-mode analysis to account for a polarized
sky observed with a polarized antenna beam. Heretofore this generalization has been neglected in the
discussion of $m$-mode analysis imaging.  At low frequencies, increasingly rapid Faraday rotation
leads to depolarization. Therefore polarization fractions are generally expected to decrease at low
frequencies (varying with ionospheric conditions). \citet{2016ApJ...830...38L} detected the presence
of diffuse polarized emission on degree angular scales with the MWA, also finding typical
depolarization ratios of $\sim0.3$ for pulsars at 154~MHz relative to 1.4~GHz, although there was a
large variance between pulsars. Even more depolarization is expected at frequencies $\le
73.152$~MHz, but crossed-dipole antennas with extremely large primary beams will naturally introduce
large polarization leakage terms at low elevations.  It is instructive to compute what impact this
will have on the unpolarized imaging process.

In order to understand the effect of polarization leakage, a point source with 10\% polarization in
Stokes-$Q$ was simulated at the location of Cas A.  The simulated visibilities were computed using
the measured beams for $x$- and $y$-dipoles. Because the amplitude of the two beams are not equal in
every direction on the sky, this introduces a direction-dependent leakage of Stokes-$Q$ into
Stokes-$I$. At 73.152~MHz, this leakage is $\lesssim5\%$ above $15^\circ$ elevation, but rapidly
rises to $\gtrsim50\%$ at lower elevations. Cas A is a circumpolar source and spends about 7 hours
every day skirting the horizon where the polarization leakage is large, so by placing the simulated
source at the location of Cas A, we are engineering a situation where the polarization leakage from
Stokes-$Q$ into Stokes-$I$ will be large. However, the impact on the unpolarized $m$-mode analysis
maps is mild, amounting to a 0.5\% error in the flux of the source with no measurable effect on the
PSF.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparisons with Other Sky Maps}\label{sec:compare}

\subsection{Guzm\'{a}n 45 MHz Map}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/guzman/guzman}
    \caption{
        This Mollweide projection map compares the fractional difference between the Guzm\'{a}n 45
        MHz map, and the OVRO-LWA map at 46.992 MHz (degraded to $5^\circ$ resolution). A positive
        value indicates regions where the OVRO-LWA map has more emission that the Guzm\'{a}n map,
        and a negative value indicates regions where the Guzm\'{a}n map has more emission than the
        OVRO-LWA map. Cas A, Cyg A, Vir A, and Tau A are masked due to the fact that they are
        subtracted from the OVRO-LWA maps but not the Guzm\'{a}n map.
    }
    \label{fig:guzman-comparison}
\end{figure*}

A direct comparison between the OVRO-LWA maps interpolated to 45 MHz and the Guzm\'{a}n 45 MHz map
\citep{2011A&A...525A.138G} can be seen in Figure~\ref{fig:guzman-comparison}. In order to make this
comparison the OVRO-LWA map was degraded to a 5$^\circ$ resolution by convolving with a Gaussian
kernel, and the Guzm\'{a}n map has had spherical harmonics with $m=0$ discarded in order to make it
consistent with the maps presented in this paper. This figure shows a $\sim20\%$ excess of emission
in the galactic plane that is consistent with the discrepancy observed by the LWA1 sky survey
\citep{2017MNRAS.469.4537D}. However, while the LWA1 sky survey has an excess of emission near the
north galactic pole, no such excess is observed in this work. Instead there is a 10\% excess of
emission near the south galactic pole. Elsewhere off the plane of the galaxy the discrepancy can be
as much as $-20\%$.

\subsection{Haslam 408 MHz Map}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/haslam/better-haslam-spectral-index}
    \caption{
        \todo{update caption}
        The contours show the spectral index between the reprocessed Haslam 408 MHz map and the
        73.152 MHz map (degraded to 56 arcmin resolution). These contours are overlaid on the same
        three-color map from Figure~\ref{fig:three-color}. The regions with shallow spectral index
        between 73.152 MHz and 408 MHz coincide with regions with large free-free absorption that is
        evident between 36.528 MHz and 73.152 MHz.
    }
    \label{fig:haslam-spectral-index}
\end{figure*}

The Haslam 408 MHz map \citep{1981A&A...100..209H, 1982A&AS...47....1H} was reprocessed by
\citet{2015MNRAS.451.4311R} to remove artifacts associated with $1/f$ noise and bright sources. In
order to compare the OVRO-LWA maps with the lower resolution Haslam map, the OVRO-LWA maps were
degraded to 56 arcmin resolution and the spectral index between the 73.152 MHz map and the 408 MHz
Haslam map was computed by means of a local T-T plot at each pixel in the maps (in order to obviate
the need for a zero-level correction to the 73.152 MHz map). The results of this computation are
shown in Figure~\ref{fig:haslam-spectral-index}. Because this comparison uses a local T-T plot to
estimate the spectral index, the angular resolution of this computation is somewhat degraded
relative to other comparisons in the literature but the spectral indices in
Figure~\ref{fig:haslam-spectral-index} are robust to zero-level errors in either map.

The primary difference between the spectral index derived between 73.152 MHz and 408 MHz
(Figure~\ref{fig:haslam-spectral-index}), and the spectral index derived between 36.528 MHz and
73.152 MHz (Figure~\ref{fig:internal-spectral-index}) is the appearance of a multitude of \ion{H}{2}
regions that dot the galactic plane in the latter. The free-free absorption from these \ion{H}{2}
regions contributes to an overall flattening of the spectral index in the galactic plane that is
also seen at higher frequencies.

%\subsection{The Global Sky Model}?
%\subsection{LWA1 Sky Survey}?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:conclusion}

In this work we presented a new imaging technique -- Tikhonov regularized $m$-mode analysis imaging
and cleaning -- for drift-scanning telescopes like the OVRO-LWA.  This technique exactly corrects
for widefield effects in interferometric imaging with a single synthesis imaging step.  We applied
Tikhonov regularized $m$-mode analysis imaging to a 28 hour dataset and generated 8 sky maps between
36.528~MHz and 73.152~MHz.  These sky maps are a substantial improvement over existing maps at these
frequencies due to their 15~arcmin angular resolution and $<600$~K thermal noise. The point source
flux scale is consistent with that defined by \citet{2012MNRAS.423L..30S} to about 5\% and large
angular scales are consistent with the work of \citet{2011A&A...525A.138G} to within 20\%.  These
maps are primarily intended to be used as part of a foreground modeling and subtraction routine for
21~cm cosmology experiments.

At frequencies above $\sim55$~MHz, the angular resolution of these maps is limited by the selection
of $l_\text{max}=1000$. Future work will increase $l_\text{max}$ to remove this restriction, as well
as include more time and bandwidth to improve the thermal noise. The usage of nighttime only data
can help mitigate dynamic range limitations from the ionosphere and also eliminate solar sidelobe
residuals. Observations could also be extended to slightly higher and lower frequencies ($\sim27$ to
$85$~MHz) to take advantage of the full frequency range of the OVRO-LWA. The higher frequencies are
particularly interesting in order to maximize the overlap with the MWA in the southern hemisphere,
which could be used to fill-in the hole around the southern celestial pole.

\acknowledgments
This work is dedicated to the memory of Professor Marjorie Corcoran, who was an influential mentor
to M.~W.~Eastwood.

\bibliographystyle{apj}
\bibliography{paper}

\end{document}